---
title: "MTCars Report"
author: "Lucy Gao"
date: "2023-08-02"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Motor Trend Car Road Tests data

The Motor Trend Car Road Tests (`mtcars`) data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

For now, we need to remove three erroneous data points: the Mazda RX4, the Valiant, and the Volvo 142E.

We print out the data below. 

```{r}
mtcars
```

## Number of gears vs. transmission 

Let's find out how many cars have different numbers of gears. 

```{r}
table(mtcars$gear)
```

Now let's explore the relationship between the number of gears and transmission. 

```{r} 
table(mtcars$am, mtcars$gear)
```

We see that no automatic cars have 5 gears, and no manual cars have 3 gears. 

## Horsepower vs. Mileage

Let's find out if there's a relationship between car horsepower vs. mileage. 

```{r} 
plot(mtcars$mpg, mtcars$hp)
```
