---
title: "Team Troubleshooting Deliverable 2"
output: github_document
---

```{r include = FALSE}
knitr::opts_chunk$set(error = TRUE)
```


## Attributions

Thanks to Icíar Fernández-Boyano for writing most of this document, and Diana Lin for her edits.
   
## Instructions

There are **24 errors** in this Rmd. Will you find them all?

+ *15 errors are marked* (at least the general area in which they are located) by ## ERROR HERE ## in the code chunk.
+ *9 errors are hiding.* 

Hint: You should be able to knit the Rmd if all code is working correctly...but this will only catch the errors in code, and not the errors in logic or instruction-following!

**Notes:**

+ Errors may be in the form of **broken code** (e.g. using " " inappropriately, which returns an error when running the code), but can also be **running code that does not follow the instructions** (e.g. the goal of the code was to filter the first 3 rows, but instead, the first 3 columns were selected). Read the Rmd commentary carefully to find these.

+ A code chunk with multiple errors, whether broken code or wrongly followed instructions, **counts as a single error for the purpose of grading**. An example:

```{r include = FALSE}
library(tidyverse)
datasets::mtcars
```

Exercise A: In the following chunk, I would like to select all columns of the mtcars dataset from `disp` to `wt` (inclusive), and then, filter for those with horsepower (`hp`) of at least 100 or greater.

```{r}
### ERROR HERE ###
mtcars %>%
  select(disp:vs) %>%
  filter(hp >= 100)
```

Notice that the instructions indicate selecting columns `disp` to `wt`, and the code reads `select(disp:vs)`. There is the error.

In addition to selecting the wrong columns, this exercise encapsulates the number 100 in between quotes - this is incorrect R syntax. In fact, if you run the code you will see that the filtering is not working.

While Exercise A has 2 requires two corrections to fix, it counts as having a *single error*, as you cannot obtain your desired output without fixing both. Therefore, errors within the *same literal R markdown code chunk* count as **one error**. For the purposes of grading, each _corrected_ literal R markdown code chunk is considered **one error** and worth **one point** each, for a total of 25 points.

Solution:

```{r}
mtcars %>%
  select(disp:wt) %>%
  filter(hp >= 100)
```

## Troubleshooting time! 

There are 24 code chunks in this Rmd with errors. Will you be the one to find them all? More than 1 error in a single chunk of code counts as a **single** error. 
  
```{r}
### ERROR HERE ###
libraries(dslabs)
libraries(singer)
libraries(tidyverse)
libraries(stringr)
libraries(gapminder)
```

[MovieLens](https://dl.acm.org/doi/10.1145/2827872) are a series of datasets widely used in education, that describe movie ratings from the MovieLens [website](https://movielens.org/). There are several MovieLens datasets, collected by the [GroupLens Research Project](https://grouplens.org/datasets/movielens/) at the University of Minnesota. Here, we load the MovieLens 100K dataset from Rafael Irizarry and Amy Gill's R package, [DSLabs](https://cran.r-project.org/web/packages/dslabs/dslabs.pdf), which contains datasets useful for data analysis practice, homework, and projects in data science courses and workshops. 

### Exercise 1

Let's have a look at the dataset! My goal is to:

+ Find out the "class" of the dataset.
+ If it isn't a tibble already, coerce it into a tibble and store it in the variable "movieLens".
+ Have a quick look at the rows and columns, using a *dplyr function*.

```{r}
### ERROR HERE ###
class(dslabs::movielens)
movieLens <- as_tibble(dslabs::movielens)
dim(movieLens)
```

### Exercise 2

Now that we've had a quick look at the dataset, it would be interesting to explore the rows (observations) in some more detail. I'd like to know how many movie entries...

+ Belong to the genre *"Drama"*?
+ Don't belong to the genre *"Drama"*?
+ Were filmed *after* the year 2000?
+ Have a rating of *4.5 or greater*?
+ Were filmed in 1999 *or* 2000?
+ Correspond to the movie *"Little Women"*?
+ Have *more than* 4.5 stars, and were released *before* 1995?

```{r}
filter(movieLens, genres == "Drama")
filter(movieLens, !genres == "Drama")
filter(movieLens, year >= 2000)
filter(movieLens, rating < 4.5)
filter(movieLens, year == 1999 | month == 2000)
filter(movieLens, title %in% "Little Women")
filter(movieLens, rating > 4.5, year < 1995)
```

### Exercise 3

While filtering for *all movies that do not belong to the genre drama* above, I noticed something interesting. I want to filter for the same thing again, this time selecting variables **title and genres first,** and then *everything else*. Hint: there is a function to select "everything else"...

```{r}
### ERROR HERE ###
movieLens %>%
  filter(!genres == "Drama") %>%
  select(title, genres, year, rating, timestamp)
```

Notice that some movies appear to have several genres, instead of a single one. If we filter for the movie "Dumbo" and select the genres column, we see that "Animation|Children|Drama|Musical" appear as genres.

```{r}
movieLens %>%
  title == "Dumbo" %>%
  select(genres)
```

This means that when we filter for "Drama", we are only filtering for those movies that are classified as *exclusively* Drama. What if we want to see all movies that *contain* "Drama" in their genres, even if they are a Romance-Drama, or a Comedy-Drama? Here is a handy, **error-free** expression that can help you do that. No errors, only for learning.

```{r}
movieLens %>%
  filter(str_detect(genres, "Drama"))
```

Now that you know that, select (and filter, where required!):

+ All colums *except* those from userId to timestamp (inclusive)
+ Columns year, title, and genres (in that order) *only* for movies classified *exclusively* as "Romance"
+ Columns rating, title, and year (in that order) followed by everything else (all other columns), only for movies *containing* the genre "Comedy"

```{r}
### ERROR HERE ###
select(movieLens, -(userId:timestamp))

movieLens %>%
  filter(genres == "Romance") %>%
  select(year, title, genres)

movieLens %>%
  filter(str_select(genres, Romance)) %>%
  select(year, title, genres)
```

### Exercise 4

Some of our variables are currently in *camelCase* (in fact, *movieLens* is in camelCase). Let's clean these two variables to be *snake_case* instead, and assign our post-rename object back to "movieLens".

```{r}
movieLens <-
movieLens %>%
  rename(user_id == userId,
         movie_id == movieId)
```

As you already know, `mutate()` defines and inserts new variables into a tibble. There is *another mystery function* that you've seen already that adds the new variable, and drops existing ones. I wanted to create an average_rating column that takes the mean(rating) across all entries, and I only want to see that variable (i.e drop all others!) but I forgot what that mystery function is. Can you remember?

```{r}
### ERROR HERE ### 
mutate(movieLens,
       average_rating = mean(rating))
```

### Exercise 5

I'm interested in arranging movies from the year 1998 by their descending rating. Then, I would like to rearrange the columns to show `rating` followed by `title`, and then followed by everything else. 

```{r}
movieLens %>%
  filter(year == 1998) %>%
  arrange(desc(rating)) %>%
  select(rating, title, everything())
```

Instead of filtering to a specific year, group the movies by year, and then arrange by ascending rating. Lastly, select only the year, title, and rating, in that order.

*Hint:* Grouped arrange needs one more argument...

```{r}
movieLens %>%
  group_by(year) %>%
  arrange(rating) %>%
  select(year:rating)
```

### Exercise 6

Alone, `tally()` is a short form of `summarise()`. `count()` is short-hand for `group_by()` and `tally()`. 

+ Each entry of the movieLens table corresponds to a movie rating by a user. Therefore, if more than one user rated the same movie, there will be several entries for the same movie. I want to find out how many times each movie has been reviewed, or in other words, how many times each movie title appears in the dataset.

```{r}
movieLens %>%
  group_by(title) %>%
  tally()
```

+ Without using `group_by()`, I want to find out how many movie reviews there have been by year.

```{r}
### ERROR HERE ###
movieLens %>%
  tally(year, sort = TRUE)
```

+ Both `count()` and `tally()` can be called repeatedly, each time adding up a level of detail. Below, I want to count the number of movie reviews by title and rating, and sort the results.

```{r}
movieLens %>%
  count(c(title, rating), sort = TRUE)
```

Not only do `count()` and `tally()` quickly allow you to count items within your dataset, `add_tally()` and `add_count()` are handy shortcuts that add an additional columns to your tibble, rather than collapsing each group. 

+ Suppose that I am interested in finding movies that have very few reviews, very few being under 5. Without using `group_by()`, add a column that counts the *number of reviews per movie title*, and then filter to those that have 5 reviews or under. Rename the column `n` to `n_reviews`.

```{r}
movieLens %>%
  add_count(title) %>%
  n <= 5 %>%
  mutate(n_reviews = n)
```

### Exercise 7

+ Calculate the mean rating by title and store it in a new variable `avg_rating`.

```{r}
movieLens %>%
  group_by(title) %>%
  summarize(avg_rating = mean(rating))
```

+ Calculate the mean rating by year and store it in a new variable `avg_rating`.

```{r}
movieLens %>%
  group_by(year) %>%
  summarize(avg_rating = mean(rating))
```

+ Using `summarize()`, find the minimum and the maximum rating by title, and store these new variables under `min_rating`, and `max_rating`, respectively. 

```{r}
### ERROR HERE ###
movieLens %>%
  mutate(min_rating = min(rating), max_rating = max(rating))
```

+ Using `group_by()` and `summarize()`, find the number of reviews per movie title, and arrange in order of descending number of reviews. Notice that we achieved the same result in the previous exercise using `tally()`.

```{r}
movieLens %>%
	group_by(title) %>%
	summarise(n_reviews = "n()") %>%
	arrange(n_reviews)
```

### Exercise 8

`across()` is a new dplyr function (`dplyr` 1.0.0) that allows you to apply a transformation to multiple variables selected with the `select()` and `rename()` syntax. For this section, we will use the `starwars` dataset, which is built into R. First, let's transform it into a tibble and store it under the variable `starWars`.

```{r}
starWars <- as_tibble(starwars)
```

+ Find the mean for all columns that are numeric.

```{r}
starWars %>%
  drop_na() %>%
  summarise(across(where(num), mean))
```

+ Find the minimum height and weight within each species.

```{r}
### ERROR HERE ###
starWars %>%
  group_by(species) %>%
  summarise(across("height", "weight", min)) %>%
  drop_na()
```

+ Find the number of distinct names and homeworlds by each species.

```{r}
### ERROR HERE ###
starWars %>%
  species %>%
  filter(n() > 0) %>%
  summarise(across(c(name, homeworld), n_distinct))
```

### Exercise 9

The `singer` package was created by Joey Bernhardt, and it contains 3 datasets that are excerpts from the Million Songs Dataset. The `songs` and the `locations` package will be used here. These two datasets have the columns `artist_name` and `title` in common.

+ The R `merge()` function automatically joins the frames by common variable names. Merge `songs` and `locations` using one of the join family of functions (`inner_join()`, `outer_join()`, `right_join()`, etc...), without specifying the key variable.

```{r include = FALSE}
songs <- singer::songs
locations <- singer::locations 
```

```{r}
### ERROR HERE ###
anti_join(songs, locations)
```

+ Join `songs` and `locations` in a way that drops all observations from `songs` that *do not* have a match in `locations`.

```{r}
### ERROR HERE ###
inner_join(songs, locations)
```

+ Combine `locations` and `songs` to find only the entries that correspond to song titles in the `locations` dataset.

```{r}
### ERROR HERE ###
left_join(songs, locations)
```

+ Merge `songs` and `locations` so that *all information* from both datasets is combined.

```{r}
### ERROR HERE ###
semi_join(songs, locations)
```

### Exercise 10

Time for reviewing... and tibbles! 

+ Create a tibble with 4 columns, `birth_year`, `name`, `birth_weight`, and `birth_location`. `birth_year` should contain years 1998 to 2005 (inclusive), `name` should contain the first 8 names in the `starWars` dataset `name` column, and `birth_weight` should take the `year` column, subtract 1995, and multiply by 0.45. `birth_location` should contain 8 randomly sampled locations in the `locations` dataset `city` column.

+ This tibble should be stored in the variable `fakeStarWars`.

```{r}
fakeStarWars <-
tibble(birth_year == "1998:2005",
       name == starWars$name[,1:8],
       birth_weight == (birth_year - 1995) * 0.45,
       birth_location == sample(locations$city, size = 8))
```

Let's look at this new tibble.

```{r}
fakeStarWars
```

+ Without assigning it to any variables, define the same tibble row-by-row with `tribble()`, with the columns in the following order: `name`, `birth_weight`, `birth_year`, `birth_location`.

```{r}
### ERROR HERE ###
tribble(
  ~name,            ~birth_year,  ~birth_weight, ~location
  "Luke Skywalker",  1.35      ,   1998        ,  Liverpool, England,
  "C-3PO"         ,  1.80      ,   1999        ,  Liverpool, England,
  "R2-D2"         ,  2.25      ,   2000        ,  Seattle, WA,
  "Darth Vader"   ,  2,70      ,   2001        ,  Liverpool, England,
  "Leia Organa"   ,  3.15      ,   2002        ,  New York, NY,
  "Owen Lars"     ,  3.60      ,   2003        ,  Seattle, WA,
  "Beru Whitesun Iars", 4.05   ,   2004        ,  Liverpool, England,
  "R5-D4"         ,  4.50      ,   2005        ,  New York, NY
)
```

Now, for the grand finale!

```{r}
gapminder_almost <- gapminder %>% filter(continent == "Europe") %>% distinct(year, .keep_all = TRUE)
```

+ Rearrange variables from the `fakeStarWars` dataset to be in the following order: `name`, `birth_year`, `birth_weight`, `birth_location`.
+ Rename `birth_year` to `year`.
+ Combine `fakeStarWars` to `gapminder_almost` by `year`, so that only those entries with a year that is present in `fakeStarWars` are merged.
+ Drop variables `country`, `continent`, `pop`, and `gdpPercap`.
+ Add a column that summarizes the number of times that `birth_location` entries are repeated, and name it "n_location".
+ Filter out locations repeated under 4 times.
+ Arrange by descending birth weight.
+ Drop any rows with NA values in any variable.
+ Join with the `starwars` dataset so that only those rows that are *present in `fakeStarWars`* are combined.
+ Calculate BMI (body mass index) by dividing weight (`mass`) by squared `height` *in meters*. Store it in a column with the name `bmi`.

```{r} 
### ERROR HERE ###
fakeStarWars %>%
  select(name, starts_with(birth)) %>%
  rename(year == birth_year) %>%
  left_join(gapminder_almost, by = "year") %>%
  select(-c(country, continent, pop:gdpPercap)) %>%
  add_count(birth_location, name = "n_location") %>%
  filter(!n_location > 4) %>%
  arrange(desc(birth_weight)) %>%
  drop_na() %>%
  inner_join("starwars") %>%
  mutate(bmi = mass / height / 100 ^ 2)
```


