---
title: "Class Meeting Schedule"
date: "Updated `r Sys.Date()`"
---

```{r, echo = FALSE, warning = FALSE}
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(dplyr))
```

Here are the topics for each class meeting. You can find the notes in the [guidebook](https://stat545guidebook.netlify.com).

Be sure to bring a laptop to every class!

There will be a TA in class to help students with the live coding exercises. 

## STAT 545A

```{r, echo = FALSE, results='asis'}
cm545 <- suppressMessages(read_csv(here("content", "cm_schedule-545a.csv")))
kable(cm545)
```


## STAT 547M

Subject to minor changes!

| Meeting No.         | Date         | Topic | 
|---------------------|--------------|-------|
| 01 | feb-25 tues    | Review of STAT 545 |
| 02 | feb-27 thurs   | Functional programming in R: Part I |
| 03 | mar-03 tues    | Functional programming in R: Part II |
| 04 | mar-05 thurs   | Functional programming in R: Part III |
| 05 | mar-10 tues    | Automate tasks and pipelines: Part I |
| 06 | mar-12 thurs   | Automate tasks and pipelines: Part II |
| 07 | mar-17 tues    | DashR: Part I |
| 08 | mar-19 thurs   | DashR: Part II |
| 09 | mar-24 tues    | DashR: styling dashboards |
| 10 | mar-26 thurs   | DashR: Deployment|
| 11 | mar-31 tues    | Creating Packages in R |
| 12 | apr-02 thurs   | TBD |