---
title: "Lecture Notes: "
subtitle: "Version 1.0.0"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```


## Learning Objectives


## Parallel Resources

This topic parallels the following resources. Take about 30 minutes to get acquainted with the following before coming to class:

- 

I suggest referring to this material when engaging with today's topic.

## Additional Resources



## Quiz



## Demonstration


<!-- ******TUESDAY'S CLASS****** -->

## Tuesday's Class `r tues_class_num`: Data Wrangling II

```{r tues-date}
cat(tues_date_pretty)
```

From today's class, students are anticipated to be able to:

+ Grouping data to compute summarized functions using `dplyr::group_by()` and `dplyr::summarize()`
+ Compute new columns using `mutate()`
+ Use `dplyr::across()` to use functions with specific columns

| Lecture | Demonstration | View |
|---------|-----------|------|
| 6-A: dplyr Part II | [.Rmd](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/demonstrations/demo6a-dplyr-part2.Rmd) | [Knitted demo](/demonstrations/demo6a-dplyr-part2) |

#### Prepare for Class

- Watch the [STAT 545 Video Series: Episode 6-A](https://youtu.be/38BA_v7BwKk)
- We recommend you read / skim through the following resources for more details
   - [Jenny's "dplyr single table verbs"](https://stat545.com/dplyr-single.html) for dplyr verbs.
   - [Hadley's "across" blog post](https://www.tidyverse.org/blog/2020/04/dplyr-1-0-0-colwise/) to learn more about `across()`

Some useful (but not mandatory) readings are:

- [dplyr vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html)

#### During Class

+ 9:30 - 9:50: In class demo, covering:
  + group_by
  + mutate
  + summarize
  + across
+ 9:50 - 10:00: Quiz.
+ 10:00 - 10:05: Break.
+ 10:05 - 10:45: Lecture worksheet.

**Complete the Class 7-A Quiz on Canvas by Wednesday, September 30th at 11:00am PST.**

| Demonstration | Rmd | View |
|---|---|---|
| 8-A: tidyr | [.Rmd](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/demonstrations/demo8a-tidyr.Rmd) | [Knitted demo](/demonstrations/demo8a-tidyr) |
