---
title: "Lecture Notes: Plotting Part II"
subtitle: "Version 1.0.0"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```


## Learning Objectives

From today's class, students are anticipated to be able to:

+ Become familiar with customizing `ggplot2` graphs.
+ Choose an appropriate plot type for Exploratory Data Analysis.
+ Understand what makes data visualizations *effective*.
+ Understand how to customize and fine-tune ggplot2 graphics to create effective visualizations.

## Parallel Resources

This topic parallels the following resources. Take about 30 minutes to get acquainted with the following before coming to class:

- [STAT 545 Episode 7-A: ggplot2 Part 2: Plotting for Exploratory Data Analysis](https://youtu.be/X8b44P5D-5w)

We suggest referring to this material when engaging with today's topic.

## Additional Resources

- [Fundamentals of Data Visualization](https://serialmentor.com/dataviz/introduction.html) by Claus Wilke.
  - Excellent guide to designing visuals that are appropriate to the data at hand. 
- [STAT545.com](https://stat545.com/effective-graphs.html) - Effective Graphics by Jenny Bryan.
- [Callingbull.org](https://callingbull.org/tools.html) by Carl T. Bergstrom and Jevin West.
  - Great resource for rethinking the "typical" way we do things. 

Here are some resources for you to investigate various aspects of graphing in more detail:

- [Visualization Analysis and Design](https://www-taylorfrancis-com.ezproxy.library.ubc.ca/books/9780429088902) by Tamara Munzner.
   - Gold-standard book for designing plots with a focus on human perception.
- [ggplot2 book](https://ggplot2-book.org/)by Hadley Wickam.
   - Goes into detail on how to use ggplot2.

## Quiz

1. True or False: Because the `dplyr::group_by()` function and ggplot2's `group` aesthetic both split a tibble into groups, we can use either one when making a plot. 
2. True or False: In principle, mapping `continent` to letters (a, b, c, ...) is an aesthetic mapping.

```{r}
gapminder %>% 
  group_by(continent) %>% 
  nrow
gapminder %>% 
  ggplot(aes(gdpPercap, lifeExp)) +
  geom_point(aes(shape = continent)) +
  scale_x_log10() +
  scale_shape_manual(values = letters[1:5])
```



## Demonstration


```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(gapminder)
knitr::opts_chunk$set(error = TRUE, fig.width = 5, fig.height = 2)
```


### Demo 1

From last class:

> Compute a weighted mean of life expectancy for different years, weighting by population.

Plot that global average life expectancy over time.

```{r}
gapminder %>% 
  group_by(year) %>% 
  summarise(wmean = sum(lifeExp * pop) / sum(pop)) %>% 
  ggplot(aes(year, wmean)) +
  geom_line()
```

### Demo 2

From last class:

> For each continent and year, how many countries have a "low" life expectancy? Let’s say life expectancy is "low" if it’s less than the median most recent life expectancy for each continent.

Plot the number of "low life expectancy" countries over time for each continent. 

```{r}
gapminder %>% 
  group_by(continent) %>% 
  mutate(median = median(lifeExp[year == 2007])) %>% 
  filter(lifeExp < median) %>% 
  count(continent, year) %>% 
  ggplot(aes(year, n)) +
  geom_line(aes(colour = continent)) +
  ylab("# of countries with low lifeExp") +
  labs(title = "Evolution over time") +
  scale_x_continuous("") +
  scale_colour_discrete("") +
  theme_minimal() +
  theme(panel.grid = element_blank())
```


### Demo 3

**(a)**

Plot GDP per capita over time, identifying each continent. 

```{r}
p3 <- ggplot(gapminder, aes(year, gdpPercap)) +
  geom_line(aes(group = country, colour = continent), alpha = 1/3) +
  facet_wrap(~ continent) +
  scale_y_log10(labels = scales::dollar_format()) +
  guides(colour = FALSE)
p3
```


**(b)**

From last class Demo 3:

> Get the maximum and minimum `gdpPercap` for all continents.

We'll also add _for each year_.  Today: modify plot (a) above so that the lines for each continent are contained within a band.

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  mutate(max = max(gdpPercap),
            min = min(gdpPercap)) %>% 
  ggplot(aes(year, (gdpPercap))) +
  geom_ribbon(aes(ymin = min, ymax = max), alpha = 0.5, fill = "gray") +
  geom_line(aes(group = country, colour = continent), alpha = 1/3) +
  # geom_ribbon(aes(ymin = quantile(gdpPercap, probs = 0.25),
  #                 ymax = quantile(gdpPercap, probs = 0.75),
  #                 group = year)) +
  facet_wrap( ~ continent) +
  scale_y_log10(labels = scales::dollar_format()) +
  guides(colour = FALSE) +
  theme_bw()
```


### Demo 4

From last class:

> Filter gapminder to all entries that have experienced a drop in life expectancy. Hint: `dplyr::lag()`.

Today: Was a drop in life expectancy just as likely for any year, or were some years more prone to see a drop than others? Is this true for all continents?

```{r}
gapminder %>% 
  group_by(country) %>% 
  arrange(year) %>% 
  mutate(inc = lifeExp - lag(lifeExp)) %>% 
  filter(inc < 0) %>% 
  ggplot(aes(year)) +
  geom_bar()
```


### Demo 5

How has the distribution of GDP per capita changed over time for each continent?

(Breakout rooms?)

```{r}
gapminder %>% 
  filter(continent != "Oceania") %>% 
  ggplot(aes(gdpPercap, factor(year))) +
  ggridges::geom_density_ridges(aes(fill = continent), alpha = 1/3) +
  scale_x_log10("GDP per capita", labels = scales::dollar_format()) +
  ylab("Year") +
  theme_minimal() +
  scale_fill_discrete("")
```


### Effective Vis

Let's go through [Principles of Figure Design](https://vincenzocoia.com/communicating_data/#28) slides, modelling Part II of [Claus Wilke's vis book](https://clauswilke.com/dataviz/).

### Demo 6

The following mock dataset marks the (x,y) position of a caribou at four time points. 

```
tribble(
  ~time, ~x, ~y,
  1, 0.3, 0.3,
  2, 0.8, 0.7,
  3, 0.5, 0.9,
  4, 0.4, 0.5
) %>% 
  ggplot(aes(x, y)) + 
  geom_line()
```

- Fix the plot below so that it shows the path of the caribou. 
- Add an arrow with `arrow = arrow()`.
- Add the `time` label with `geom_text()`.

```{r}
tribble(
  ~time, ~x, ~y,
  1, 0.3, 0.3,
  2, 0.8, 0.7,
  3, 0.5, 0.9,
  4, 0.4, 0.5
) %>% 
  ggplot(aes(x, y)) + 
  geom_path(arrow = arrow()) +
  geom_text(aes(label = time), nudge_x = 0.02) +
  theme_void()
```

### Demo 7

- Change the following plot so that it shows _proportion_ on the y-axis, not count.
- Change the x-axis so that it doesn't appear to be continuous.
- Put the bars for transmission side-by-side with their own colour.
- Capitalize the legend title.

```
mtcars %>% 
  mutate(transmission = if_else(am == 0, "automatic", "manual")) %>% 
  ggplot(aes(cyl)) +
  geom_bar(aes(colour = transmission))
```

```{r, fig.width = 5, fig.height = 2}
mtcars %>% 
  mutate(transmission = if_else(am == 0, "automatic", "manual")) %>% 
  ggplot(aes(factor(cyl))) +
  geom_bar(aes(y = ..prop.., fill = transmission, group = transmission), 
           position = "dodge") +
  scale_fill_discrete("Transmission") +
  labs(y = "Proportion", x = "Cylinders") +
  theme_minimal()
```

### Demo 8

Here's the number of people having a certain hair colour from a sample of 592 people:

```{r}
(hair <- as_tibble(HairEyeColor) %>% 
  count(Hair, wt = n))
```

Fix the following bar plot so that it shows these counts.

```
ggplot(hair, aes(Hair, n)) +
  geom_bar()
```

```{r}
ggplot(hair, aes(Hair, n)) +
  geom_col()
```

### Demo 9

Here's the number of people having a certain hair and eye colour from a sample of 592 people:

```{r}
(hair_eye <- as_tibble(HairEyeColor) %>% 
  count(Hair, Eye, wt = n))
```

### Demo 10
 
Fix the following plot so that it shows a filled-in square for each combination. 

```
ggplot(hair_eye, aes(Hair, Eye)) +
  geom_point(aes(colour = n))
```

```{r}
ggplot(hair_eye, aes(Hair, Eye)) +
  geom_bin2d(aes(fill = n)) +
  theme_minimal()
```

