---
title: "Week 1-B Lecture Notes: R Functions for Data Analysis"
subtitle: "**Development** Version 0.9000"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
week_num <- 8  # 1 through 14
#-------------
source(here::here("content", "dates.R"))
library(stringr)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "asis")
letter <- if (week_num <= 7) "A" else "B"
tues_class_num <- paste0(str_pad(week_num * 2 - 2, width = 2, pad = 0), letter)
tues_date <- lec0 + days((week_num - 1) * 7)
thurs_date <- tues_date + days(2)
```


```{r tues-date}
cat(format(tues_date, "%A, %b %d"), "-", 
    format(thurs_date, "%A, %b %d"))
```

## Learning Objectives

Key topics this week:

- Start getting a sense of when to make a function in a data analysis (we will build on this next week).
- Workflow for building a function: start interactively, wrap it as a function. `return()`. Argument names. 
- Fortify a function: 
  - generalize the function and use smart defaults; `NA` handling, and ellipses package [https://ellipsis.r-lib.org/](https://ellipsis.r-lib.org/)
  - provide useful error messages; sidebar: `if` statements
  - Unit tests, and (sidebar) assertions
- Data masking in a function.

Possible topics this week:

- `try()` and `withTimeout()`.
- environments and scoping. Perhaps `search()`; `with()`.

## Prepare for Class

Since this is the first week of Part B, I won't be anticipating you to digest material beforehand. Plus, the video for this week won't come until later in the week.

Instead, I suggest digesting this content throughout the week:

- Read [stat545.com Functions, Parts 1-3](https://stat545.com/functions-part1.html)
- Watch STAT 545 Episode 1-B video series (I will announce when it's available) 

## Quiz 1-B

Consider the following code:

```
x <- 1
f <- function(t) {
  x <- -1
  if (x > 0) return(t)
  x * t + 1
  x * t
}
f(1)
```

1. True or False: the output of `f(1)` is `-1`. 
2. True or False: after running the above code, `x` is `-1`. 
