---
title: "Lecture Notes: Tibble Joins"
subtitle: "Version 1.0.0"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```


## Learning Objectives

From today's class, students are anticipated to be able to:

- Recognize how to manipulate data through a variety of tibble joins such as:
  - left_join
  - right_join
  - full_join
  - anti_join
- Perform set operations on data

## Parallel Resources

This topic parallels the following resources. Take about 30 minutes to get acquainted with the following before coming to class:

- [STAT 545 Episode 11-A: Tibble Joins with dplyr](https://youtu.be/YAdX9MVRY1c)
- "Two-table verbs" [vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/two-table.html)

We suggest referring to this material when engaging with today's topic.

## Additional Resources

- [Jenny's Join Cheatsheet](https://stat545.com/join-cheatsheet.html) for a quick reference to joins.
- [Relational Data chapter](https://r4ds.had.co.nz/relational-data.html) in "R for Data Science".
- [dplyr cheatsheet](https://rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)

## Quiz

_No quiz today_

## Demonstration


```{r setup, message=FALSE, warning=FALSE}
library(palmerpenguins)
library(tidyverse)
```

This is a quick demo of joining with the penguins tibble that we saw in class.

Recall the `penguins` tibble:

```{r}
penguins
```

There are only 3 islands:

```{r}
penguins %>% 
  pull(island) %>% 
  levels()
```

Suppose we think land mass of the islands has something to do with body mass. We'd want to add land mass to the tibble. Here's a tibble of (fake) land mass values for islands in the area (I added another fake one, too):

```{r}
(mass <- tribble(
  ~ island, ~ mass,
  "Biscoe"    , 500,
  "Dream"     , 600,
  "Foo Island", 300,
  "Torgersen" , 10000
))
```

Add them to the original tibble:

```{r}
(big <- left_join(penguins, mass, by = "island"))
```

By the way, `left_join()` is probably the most useful and the most used join (in my opinion).

Notice that Foo Island hasn't changed anything about our new tibble -- there are still three islands. Let's see: 

(Note: it seems `island` is no longer a factor, so we'll use `unique()` instead of `levels()` to get all unique islands):

```{r}
big %>% 
  pull(island) %>% 
  unique()
```

Now suppose you want to do the opposite: you only have one big tibble, and you want to store island land mass in a new tibble to save storage space on your computer. We have two options:

```{r}
big %>%
  group_by(island) %>%
  summarise(mass = unique(mass))
```


```{r}
big %>%
  group_by(island) %>%
  summarise(mass = mass[1])
```

The second example is less desirable: what if -- for whatever reason -- there's _more than one_ land mass for an island? `mass[1]` wouldn't catch that, but `unique()` would. Better to use `unique()` so that you'll be made aware that there's more than one land mass.

