---
title: "Week 4-B Lecture Notes: Automation"
subtitle: "**Development** Version 0.9000"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
week_num <- 11  # 1 through 14
#-------------
source(here::here("content", "dates.R"))
library(stringr)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "asis")
letter <- if (week_num <= 7) "A" else "B"
tues_class_num <- paste0(str_pad(week_num * 2 - 2, width = 2, pad = 0), letter)
tues_date <- lec0 + days((week_num - 1) * 7)
thurs_date <- tues_date + days(2)
```


```{r tues-date}
cat(format(tues_date, "%A, %b %d"), "-", 
    format(thurs_date, "%A, %b %d"))
```

## Learning Objectives

This week aims to provide you with a foundation for making a dashboard using R shiny. From today's class, students are anticipated to be able to:

- develop basic shiny apps from scratch,
- deploy a shiny app,
- create interactive Shiny R Markdown documents, and
- research more advanced shiny techniques.

And possibly:

- Make an interactive plot with plotly
- Make a shiny document (embed shiny features within an Rmd file)


## Prepare for Class

Read/watch one of the following two resources before class:

Prep for this week is also optional. We'll be following along with [Dean Attali's shiny tutorial](https://deanattali.com/blog/building-shiny-apps-tutorial/) this week.

## Additional Resources

General:

- STAT 545 Episode 4-B
- ["Mastering Shiny"](https://mastering-shiny.org/) by Hadley Wickham
- The [official shiny site](https://shiny.rstudio.com/) is also very useful. It has tutorials, a gallery, and other goodies.

Specific tasks:

- For deploying shiny apps, check out [https://www.shinyapps.io/](https://www.shinyapps.io/)
- For shiny documents (= R Markdown + shiny), see [Chapter 19 of Yihui’s R Markdown book](https://bookdown.org/yihui/rmarkdown/shiny-documents.html).
- For testing your shiny app, check out the [shinytest](https://rstudio.github.io/shinytest/) and [reactlog](https://github.com/rstudio/reactlog) packages. 

Other dashboard tools besides shiny (but not in R):

- With python: plotly dash. Check out the [main website](https://plotly.com/dash/), or [this Medium post](https://medium.com/plotly/introducing-dash-5ecf7191b503) introducing the tool.
- With javascript: [D3](https://d3js.org/), a tremendously powerful tool with a steep learning curve (esp. if you don't know javascript).