---
title: "Lecture Notes: R Packages for Data Analysis"
subtitle: "Version 1.0"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
week_num <- 10  # 1 through 14
#-------------
source(here::here("content", "dates.R"))
library(stringr)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "asis")
letter <- if (week_num <= 7) "A" else "B"
tues_class_num <- paste0(str_pad(week_num * 2 - 2, width = 2, pad = 0), letter)
tues_date <- lec0 + days((week_num - 1) * 7)
thurs_date <- tues_date + days(2)
```


```{r tues-date}
cat(format(tues_date, "%A, %b %d"), "-", 
    format(thurs_date, "%A, %b %d"))
```


Why make an R package? As mentioned in Week 1-B, your analysis will probably benefit from homemade functions: making functions forces you to think about your analysis in terms of its key computational parts, and makes for robust and readable code. Here are a few benefits that result by bundling these functions into an R package:

- Easily access function documentation
- Easy to share
- Built-in checks to ensure your package is working
- Provides a template for organizing your work

The alternative is keeping the functions stored in separate files, and `source()`ing them into your analysis scripts, but this can become unwieldy. 

## Learning Objectives

*We will continue our discussion on purrr on Tuesday*

From today's class, students are anticipated to be able to build a basic R package, especially using the `devtools` package. 

- Write a DESCRIPTION file
- Carefully curate package dependencies
- Document functions and data using roxygen2 comments and tags
- Include tests with testthat in accordance with the R package infrastructure.
- Add a license
- Update an R package via semantic versioning, NEWS, changelog.
- Develop and build informative vignettes and a package README.
- Add a code of conduct

And possibly:

- Make a package website with pkgdown.
- Make a code coverage report and page with covr.
- Add badges to the README.
- Build an S3 object oriented package


## Prepare for Class

Prep for this week is also optional. We'll be following along with ["The Whole Game" Chapter of "R packages"](https://r-pkgs.org/whole-game.html) this week.


## Additional Resources

- STAT 545 Episode 3-B
- ["R Packages, ed. 2", by Hadley Wickham and Jenny Bryan](https://r-pkgs.org/index.html)
  - A digestible guide to making R packages, especially using devtools and friends to help.
- [Writing R Extensions](https://cran.r-project.org/doc/manuals/R-exts.html)
  - The official and comprehensive (but dry) resource for writing R packages.
- [Package development cheat sheet](https://rawgit.com/rstudio/cheatsheets/master/package-development.pdf)
- ["Advanced R" Chapter 13: S3](https://adv-r.hadley.nz/s3.html)
  - A digestible overview to S3 Object Oriented objects, like `mean()`, and how to make them. 

## Quizzes

It might help to browse a completed package directory to answer these questions, like the [powers package](https://github.com/vincenzocoia/powers).

**Quiz 6-B**

1. True or False: The mandatory files in an R package are a DESCRIPTION file, a NAMESPACE file, and functions in the R/ directory -- everything else is just helpful.
2. So far we've been using devtools functions in the console (instead of saved in a script). True or False: if you're making a bigger R package, you should start putting those devtools functions in one or more R scripts.




## Demonstration

For this topic, we'll be making an R package similar to the toy [powers package](https://github.com/vincenzocoia/powers). We'll mostly be following along with ["The Whole Game" Chapter of "R packages"](https://r-pkgs.org/whole-game.html). Here are the actions digested from that chapter, with the corresponding section number for more info. 

First, make a minimal viable product:

- Load the `devtools` package in the console (do this every time you go to work on your package).
- 2.4: Run `create_package()` in the console (all devtools functions should be written in the console).
  - Better defaults than going through the File menu.
- 2.5: `use_git()`
  - Prefer to start your project on GitHub? Or locally? Either way, useful instructions for what to do can be found [in "Happy Git with R" Part III](https://happygitwithr.com/usage-intro.html).
- 2.6 and 2.7: Make a new R script using `use_r()`, and write a function there: start with `square()`.
- 2.8: Test drive after using `load_all()`.
- 2.9: Check that the package is intact: either run `check()` or click "Check".
- 2.10 and 2.11: Edit the DESCRIPTION file. For the license, `use_mit_license("Your Name")`. 
- 2.12: Document the function: "Code > Insert roxygen skeleton", then run `document()`.
- 2.14: Install and Restart, or run `install()`. Try loading the package and using it!

Now, expand:

- 2.15: `use_testthat()` for your package, and make an R script containing tests 
  - Check all tests with `test()`. This also happens with `check()` or the "Check" button.
- 2.16: Expand `square()` so that it can take either a vector or list, using purrr's `map_dbl()` function. Indicate dependency with `use_package()`. 
- 2.17: Add a package README with `use_readme_rmd()`, and render it every time with `build_readme()`.
  - Once you successfully sync your package to GitHub, other people can install the package with `devtools::install_github("your_username/package_name")` -- this should go in the README.
- 11.1: Make a vignette with `use_vignette()`. Build all vignettes with `build_vignettes()`.
- 14.1: Include data with the R package with `use_data(R_OBJECT_HERE)`. Then document _a string of its name_ in a new R script using [a different collection of roxygen tags](https://r-pkgs.org/data.html#documenting-data).

Release your package (Chapter 20):

- 20.5.3: Make a NEWS.md file with `use_news_md()` and add the main development notes.  
- 20.1: Choose a version number; put in description file, and tag a release on GitHub.
  - 20.7: You should also prepare your package for the next version -- I suggest doing this on a new branch.