---
title: "Lecture Notes: "
subtitle: "Version 1.0.0"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```


## Learning Objectives


## Parallel Resources

This topic parallels the following resources. Take about 30 minutes to get acquainted with the following before coming to class:

- 

I suggest referring to this material when engaging with today's topic.

## Additional Resources



## Quiz



## Demonstration


-----

<!-- ******THURSDAY'S CLASS****** -->

## Thursday's Class `r thurs_class_num`: Tibble Joins

```{r thurs-date}
cat(thurs_date_pretty)
```

From today's class, students are anticipated to be able to:

- Recognize how to manipulate data through a variety of tibble joins such as:
  - left_join
  - right_join
  - full_join
  - anti_join
- Perform set operations on data

| Lecture | Demonstration | View |
|---------|-----------|------|
| 11-A: tibble joins | [.Rmd](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/demonstrations/demo11a-joins.Rmd) | [Knitted demo](/demonstrations/demo11a-joins) |

#### Prepare for Class

- Watch the [STAT 545 Video Series: Class `r thurs_class_num`](https://youtu.be/YAdX9MVRY1c)
- Read the "two-table verbs"'s [vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/two-table.html)

Some useful (but not mandatory) readings are:

- [Jenny's Join Cheatsheet](https://stat545.com/join-cheatsheet.html) for a quick reference to joins.
- [Relational Data chapter](https://r4ds.had.co.nz/relational-data.html) in "R for Data Science".
- [dplyr cheatsheet](https://rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)

#### During Class

- In class demo, covering:
- [__Mutating joins__](https://r4ds.had.co.nz/relational-data.html#mutating-joins) add new columns to the "original" tibble.
- [__Filtering joins__](https://r4ds.had.co.nz/relational-data.html#filtering-joins) filter the "original" tibble's rows.
- [__Set operations__](https://r4ds.had.co.nz/relational-data.html#set-operations) work as if each row is an element in a set. 
- __Binding__ stacks tables on top of or beside each other, with `bind_rows()` and `bind_cols()`.

- Quiz: **No quiz** ~~Complete the Class `r thurs_class_num` Quiz on Canvas by Friday, `r fri_date_pretty` at 11:00am PST.~~
