---
title: "Week 3-B Lecture Notes: R Packages for Data Analysis"
subtitle: "**Development** Version 0.9000"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
week_num <- 10  # 1 through 14
#-------------
source(here::here("content", "dates.R"))
library(stringr)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "asis")
letter <- if (week_num <= 7) "A" else "B"
tues_class_num <- paste0(str_pad(week_num * 2 - 2, width = 2, pad = 0), letter)
tues_date <- lec0 + days((week_num - 1) * 7)
thurs_date <- tues_date + days(2)
```


```{r tues-date}
cat(format(tues_date, "%A, %b %d"), "-", 
    format(thurs_date, "%A, %b %d"))
```

## Learning Objectives

From today's class, students are anticipated to be able to build a basic R package, especially using the `devtools` package. 

- Write a DESCRIPTION file
- Carefully curate package dependencies
- Document functions and data using roxygen2 comments and tags
- Include tests with testthat in accordance with the R package infrastructure.
- Add a license
- Update an R package via semantic versioning, NEWS, changelog.
- Develop and build informative vignettes and a package README.
- Add a code of conduct

And possibly:

- Make a package website with pkgdown.
- Make a code coverage report and page with covr.
- Add badges to the README.
- Build an S3 object oriented package


## Prepare for Class

Prep for this week is also optional. We'll be following along with ["The Whole Game" Chapter of "R packages"](https://r-pkgs.org/whole-game.html) this week.


## Additional Resources

- STAT 545 Episode 3-B
- ["R Packages, ed. 2", by Hadley Wickham and Jenny Bryan](https://r-pkgs.org/index.html)
- [Writing R Extensions](https://cran.r-project.org/doc/manuals/R-exts.html): The official and comprehensive (but dry) resource for writing R packages.
