---
title: "Week 2-B Lecture Notes: List Columns"
subtitle: "**Development** Version 0.9000"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
week_num <- 9  # 1 through 14
#-------------
source(here::here("content", "dates.R"))
library(stringr)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "asis")
letter <- if (week_num <= 7) "A" else "B"
tues_class_num <- paste0(str_pad(week_num * 2 - 2, width = 2, pad = 0), letter)
tues_date <- lec0 + days((week_num - 1) * 7)
thurs_date <- tues_date + days(2)
```


```{r tues-date}
cat(format(tues_date, "%A, %b %d"), "-", 
    format(thurs_date, "%A, %b %d"))
```

## Learning Objectives

From today's class, students are anticipated to be able to:




## Prepare for Class

Read/watch one of the following two resources before class:

- Watch STAT 545 Episode
- 

## Resources



## Outline

Nesting, list columns, rectangling, purrr. Revisit `across()`?

- Draw from Jenny's https://jennybc.github.io/purrr-tutorial/ls01_map-name-position-shortcuts.html
- Possible new resource for nested data: [https://tidyr.tidyverse.org/articles/nest.html](https://tidyr.tidyverse.org/articles/nest.html)
- rectangling may be most useful for JSON files -- not worth digging into deeply?
- Use cases (as per [https://r4ds.had.co.nz/many-models.html#creating-list-columns](https://r4ds.had.co.nz/many-models.html#creating-list-columns))
    - Nesting: `nest()` and `unnest()`
    - rectangling for JSON data: `unnest_wider()`, `unnest_longer()`, and `hoist()`
    - purrr list outputs with `mutate()`
- Draw examples from my analyses over the past year or so.
- `do.call()`