---
title: "Lecture Notes: "
subtitle: "Version 1.0.0"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```


## Learning Objectives


## Parallel Resources

This topic parallels the following resources. Take about 30 minutes to get acquainted with the following before coming to class:

- 

I suggest referring to this material when engaging with today's topic.

## Additional Resources



## Quiz



## Demonstration




<!-- ******THURSDAY'S CLASS****** -->

## Thursday's Class `r thurs_class_num`: ggplot2 II 

```{r thurs-date}
cat(thurs_date_pretty)
```

From today's class, students are anticipated to be able to:

+ Become familiar with customizing `ggplot2` graphs 
+ Choose an appropriate plot type for Exploratory Data Analysis
+ Understand what makes data visualizations *effective*


#### Prepare for Class

- Watch the [STAT 545 Video Series: Episode 7-A](link)

Some useful (but not mandatory) readings related to today are:

- [Fundamentals of Data Visualization](https://serialmentor.com/dataviz/introduction.html) by Claus Wilke.
  - Excellent guide to designing visuals that are appropriate to the data at hand. 
- [STAT545.com](https://stat545.com/effective-graphs.html) - Effective Graphics by Jenny Bryan.
- [Callingbull.org](https://callingbull.org/tools.html) by Carl T. Bergstrom and Jevin West.
  - Great resource for rethinking the "typical" way we do things. 

#### During Class

+ 9:30 - 9:50: In class demo, covering:
  + customizing ggplot2 graphs
  + effective visualizations
+ 9:50 - 10:00: Quiz.
+ 10:00 - 10:05: Break.
+ 10:05 - 10:45: Lecture worksheet.

**Complete the Class 7-A Quiz on Canvas by Friday, October 2nd at 11:00am PST.**

| Lecture | Demonstration | View |
|---------|-----------|------|
| 7-A: ggplot2 Part II | [.Rmd](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/demonstrations/demo7a-ggplot2-part2.Rmd) | [Knitted demo](/demonstrations/demo7a-ggplot2-part2) |



#### Deeper Dive

Here are some resources for you to investigate various aspects of graphing in more detail:

- [Visualization Analysis and Design](https://www-taylorfrancis-com.ezproxy.library.ubc.ca/books/9780429088902) by Tamara Munzner.
   - Gold-standard book for designing plots with a focus on human perception.
- [ggplot2 book](https://ggplot2-book.org/)by Hadley Wickam.
   - Goes into detail on how to use ggplot2
