---
title: "Week 1-B Lecture Notes: R Functions for Data Analysis"
subtitle: "**Development** Version 0.9000"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
week_num <- 8  # 1 through 14
#-------------
source(here::here("content", "dates.R"))
library(stringr)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "asis")
letter <- if (week_num <= 7) "A" else "B"
tues_class_num <- paste0(str_pad(week_num * 2 - 2, width = 2, pad = 0), letter)
tues_date <- lec0 + days((week_num - 1) * 7)
thurs_date <- tues_date + days(2)
```


```{r tues-date}
cat(format(tues_date, "%A, %b %d"), "-", 
    format(thurs_date, "%A, %b %d"))
```

## Learning Objectives

From today's class, students are anticipated to be able to:




## Prepare for Class

Read/watch one of the following two resources before class (these will suffice for the week):

- Watch STAT 545 Episode 1-B
- Read stat545.com Functions, Parts 1-3

## Resources



## Outline

- Many of the sections are on the philosophy of function making in an analysis.
- start interactively, then put it in a function.
- fail loudly and quickly: giving useful error messages
- assertions
- Generalizing your function; use arguments and smart defaults.
- `return()`
- argument names: be intentional
- `NA`s, and their handling in the function. Propagate? `na.rm`
- ellipses, and ellipses package [https://ellipsis.r-lib.org/](https://ellipsis.r-lib.org/)
- testthat
- `try()`
- `withTimeout()`?
- tidy eval? Could be a big rabbit hole.
- environments and scoping. Perhaps `search()`; `with()`.
- S3 objects, generic functions, and methods?


Grab from:

- [Guidebook 2018](https://github.com/UBC-STAT/stat-545-guidebook/blob/master/cm202.Rmd)
- [Worksheet 2018](https://github.com/UBC-STAT/stat-545-guidebook/blob/pre-leave/tutorials/cm101-exercise.Rmd)
