---
title: "STAT 545B Course syllabus: 2024/2025"
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
  # pdf_document:
  #   toc: true
  #   toc_depth: 2
  #   number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(lubridate)
source(here::here("content", "dates.R"))
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

How to make a clean and modern data analysis, Part II.

**Website**: <https://stat545.stat.ubc.ca/>

**Duration**: `r format(dates_545b[1], "%a, %b %d")` -
`r format(dates_545b[14], "%a, %b %d")`

-   How to write functions and work with R in a functional style
-   Advanced computation on data: manipulating strings, and functional
    programming with purrr
-   Distribute data and code via an R package.
-   Interactive pages, apps, and graphics with
    [Shiny](http://shiny.rstudio.com).

# Teaching Team

Instructor: Dr. Katie Burak 

Teaching Assistants:

-   Asfar Lathif
-   Erick Isaac Navarro Delgado
-   Omar Tariq 

# Meeting Schedule

Tuesdays and Thursdays 0900-1030 PT. 

**Be sure to bring a CHARGED laptop to every class!**

There will always be TA's available during class to help students with the live coding exercises.

```{r, results = "asis"}
here::here("content", "lecture_schedule-545b.csv") %>% 
  read_csv() %>% 
  mutate(Weekday = wday(dates_545b[-7], label = TRUE),
         Date = dates_545b_pretty_noyear[-7]) %>% 
  select(Class, Weekday, Date, Topic) %>% 
  knitr::kable()
```

<!-- __Non COVID-19 years__: Keep things similar, except synchronous part to be held in-person. -->

<!-- - Lecture done in-person, but with the same structure.  -->

<!--     - Live demo: Still use chat (Slack) so that students can ask questions if they don't want to speak up; still use TA to moderate these. -->

<!--     - Lab part: allow breakout groups to happen just by people's physical presence; 2 teaching team members check in person-by-person, while one keeps watch for hands. -->

# Deliverables

| Deliverable    | Percent Grade | Description |
|----------------|---------------|-------------------------|
| Class worksheets | 15 | Autograded walkthroughs to guide student learning. |
| Assignments | 85 | e.g. adding to a data analysis, making a small R package, making a Shiny dashboard | 


More details can be found [on the course
dashboard](https://stat545.stat.ubc.ca/course).

# Auditing Students

Auditing students are expected to complete all deliverables. The
difference between enrolling for credit is that auditing students are
graded on each assignment on a pass/fail basis.

# Privacy

## Slack

STAT 545 uses Slack for course communications. Note that the messages
sent on Slack are stored on servers outside of Canada.

## GitHub.com

STAT 545 asks students to work on github.com. Please produce work
knowing that the material you put on GitHub will be stored on servers
outside of Canada.

# Policies

In addition to [UBC's Campus-wide Policies and
Regulations](http://www.calendar.ubc.ca/vancouver/?tree=3,0,0,0), STAT
545A and STAT 545B adopt the following policies.

## Communications

The teaching team can't guarantee that they will be able to respond to
student messages outside of typical workday hours (0900-1700 PT). So,
please be mindful of a **17:00 PT cutoff on Fridays** when asking
assignment-related questions.

Please read [this](/slack_communication) before messaging the teaching
team.

## Late Policy

A late submission is defined as any work submitted
after the deadline. For a late submission, the student will receive a
warning for the first occurrence, and will receive a
grade of 0 for subsequent occurrences.

## Academic Concession

UBC no longer requires a doctor's note (or supporting documentation) for
[academic
concession](http://www.calendar.ubc.ca/vancouver/index.cfm?tree=3,48,0,0).
A self-declaration will suffice -- [here](/concession_template.pdf) is a
template you can use. Please submit this to the instructor.

Examples of "conflicting responsibility" are conference travel and time-sensitive field work.

If you arrange to have an assignment submitted late, you may not be able
to receive feedback from your peers.

## Plagiarism

Plagiarism, which is intellectual theft, occurs where an individual
submits or presents the oral or written work of another person as his or
her own and can include:

-   multiple students submitting the same response
-   copying from sources without citing them
-   copying verbatim (word-for-word) from source and citing, but failing
    to make it explicit that this is a quotation (quotations should be
    used only rarely, if at all)

Students are responsible for ensuring that
any work submitted does not constitute plagiarism. Students who are in
any doubt as to what constitutes plagiarism should consult their
instructor before handing in any assignments.

For more information see the [UBC Academic Misconduct
policies](http://www.calendar.ubc.ca/vancouver/index.cfm?tree=3,54,111,959).

### Code Plagiarism

Students must correctly cite any code that has been authored by someone
else or by the student themselves for other assignments. Cases of code
plagiarism may include, but are not limited to:

-   the reproduction (copying and pasting) of code with none or minimal
    reformatting (e.g., changing the name of the variables)
-   the translation of an algorithm or a script from a language to
    another
-   the generation of code by automatic code-generations software

An "adequate acknowledgement" requires a detailed identification of the
(parts of the) code reused and a full citation of the original source
code that has been reused.

## UBC's Policies and Resources to Support Student Success

UBC provides resources to support student learning and to maintain
healthy lifestyles but recognizes that sometimes crises arise and so
there are additional resources to access including those for survivors
of sexual violence. UBC values respect for the person and ideas of all
members of the academic community. Harassment and discrimination are not
tolerated nor is suppression of academic freedom. UBC provides
appropriate accommodation for students with disabilities and for
religious, spiritual and cultural observances. UBC values academic
honesty and students ae expected to acknowledge the ideas generated by
others and to uphold the highest academic standards in all of their
actions. Details of the policies and how to access support are available
[here](https://senate.ubc.ca/policies-resources-support-student-success).

## Your personal health

If you have not yet had a chance to get vaccinated
against COVID-19 or the flu, vaccines are available to you, free, and on campus. The higher the rate of
vaccination in our community overall, the lower the chance of spreading disease. You are an important part of the UBC community. Please
arrange to get vaccinated if you have not already done so.

**If you're sick, it's important that you stay home -- no matter what
you think you may be sick with (e.g., cold, flu, other).** Your precautions will help reduce risk and keep everyone safer. The structure of this class is intended to provide flexibility so that you can prioritize your health and still be able to succeed. 

**If you do miss class because of illness:**

-   Consult the class resources on the course website,
    <https://stat545.stat.ubc.ca/>.
-   Come to office hours on Zoom.
-   Use Slack to carry out discussions.

For additional information about academic concessions, see the UBC
policy here:
<http://www.calendar.ubc.ca/vancouver/index.cfm?tree=3,329,0,0>

## Instructor health

**If I (the instructor) am sick:** If I am ill, then I will not come to class. If that
happens, here's what you can expect:

-   If I am well enough to teach, I will conduct virtual lectures
    through Zoom until I am well. If this happens, you will be tagged in
    an announcement via Slack with information. You can anticipate that
    this would very likely be a last minute announcement. Our classroom
    will still be available for you to sit and attend an online session,
    although it is recommended that you bring headphones.
-   If I am not well enough to teach, it is possible that one or more
    teaching assistants will take my place. But if not, we will either
    try to make up for lost time, make new resources to aid in your
    learning, or make accommodations regarding the assignments.
