---
title: "assignment_02"
author: "YOUR_NAME"
date: "Updated `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 02

Below are 6 exercises to help you practice writing unit tests using the `testthat` package in R. You will be writing unit tests for the vectorized functions you created in assignment 1.
This assignment was designed to be completed in about 45 minutes and is deliberately designed to be shorter to accommodate the team project.

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(testthat)
mtcars_original <- mtcars
```

### Task 1: Test function inputs and outputs

#### 1.1 - Below is a function that computes the mean of every column in a dataframe. Use the appropriate `testhat` function to test that the function input is a dataframe.

```{r prob1}
output <- function(df){
  return(map(df, mean))
  }
```

Your solution to add a unit test to the code above (using a `testthat` function):

```{r soln1}
# BEGIN SOLUTION
output <- function(df){
  
  test_that('df must be a dataframe', {
  expect_true(is.data.frame(df))
  })
  
  return(map(df, mean))
}
# END SOLUTION
```

```{r soln1 check fail, error=TRUE}
# To check you wrote the unit test properly, this code should throw an error
output(1)
```

```{r soln1 check success, error=TRUE}
# To check you wrote the unit test properly, this code should run
output(mtcars)
```

#### 1.2 - Below is a vectorized function that divides the values in each column of the `mtcars` dataset by the maximum in that column. Use the appropriate `testthat` function to check that the first value in the output is equal to 0.619469 (hint: to retrieve the first value `output[[1]][1]`) (hint: set `tolerance=1e-5`).

```{r prob2}
output <- map(mtcars, ~./max(.))
output
```

Your solution to add a unit test to the code above (using a `testthat` function):

```{r soln2}
# BEGIN SOLUTION
test_that('first value should be 0.619469', {
  expect_equal(output[[1]][1], 0.619469, tolerance=1e-5)
  })
# END SOLUTION
```

#### 1.3 - Below is a vectorized function that creates a ggplot with fuel efficiency on the x-axis and horsepower on the y-axis for three cylinder levels (4, 6, 8). Use the appropriate `testthat` function to check that the number of plots being outputted is equal to 3.

```{r prob3}
cylinders <- sort(unique(mtcars_original[['cyl']]))
output <- map(1:length(cylinders), ~ mtcars_original %>% 
    filter(cyl == cylinders[.x]) %>%
    ggplot() + 
    theme_bw() +
    geom_point(aes(x = mpg, y = hp) ) + 
    labs(x = 'Fuel efficiency (mpg)',
         y = 'Horsepower (hp)') + 
    ggtitle(glue::glue("Horsepower and Fuel efficiency for ", cylinders[.x], " cylinders")))
output
```

Your solution to add a unit test to the code above (using a `testthat` function):

```{r soln3}
# ADD SOLUTION HERE
# BEGIN SOLUTION
test_that('code should output 3 plots', {
  expect_equal(length(output), 3)
  })
#END SOLUTION
```

### Task 2: Writing your own unit tests

Some ideas for writing unit tests include checking if an input or output is the correct length, type, value, etc.

#### 2.1  - Below is sample code that computes the number of unique values in each column of `mtcars` in a vectorized way.  Use the appropriate `testthat` function to check the output of this code.

```{r prob4}
output <- map(mtcars, unique) %>% 
  map_dbl(length)
output
```

Your solution to add a unit test to the code above (using a `testthat` function):

```{r soln4}
# ADD SOLUTION HERE
# BEGIN SOLUTION
test_that('length of output should be 11', {
  expect_equal(length(output), 11)
  })
# END SOLUTION
```

#### 2.2 - Below is sample code that takes input from various columns in the `diamonds` dataset and outputs a string containing information from the input. Use the appropriate `testthat` function to write 2 tests: 1 test to check the input of the `output` function and 1 test to check the ouput of the `output` function .

```{r, prob5, echo = TRUE}
dmonds <- diamonds %>% 
  slice(1:4)
list_of_things <- list(seq(nrow(dmonds)), dmonds$price, dmonds$carat)
output <- function(input_list){
  pmap_chr(input_list,
           ~paste0("Diamond #", ..1 ,
                   " sold for $",..2,
                   " and was ",..3, " carats"))
}
output(list_of_things)
```

Your solution to add a unit test to the code above (using a `testthat` function):

```{r soln5}
# ADD SOLUTION HERE
# BEGIN SOLUTION
test_that('list_of_things must be a list', {
  expect_true(is.list(list_of_things))
  })
test_that('function should output 4 strings', {
  expect_true(is.character(output(list_of_things)))
  expect_equal(length(output(list_of_things)), 4)
  })
# END SOLUTION
```

#### 2.3 - Below is sample code that splits the iris dataset into three species, and saves each dataframe as a separate file. Use the appropriate `testthat` function to check that these files were successfully created. Bonus points if you use the appropriate `purrr` function to check if all the files were successfully created. (hint: `file.exists(here::here(glue::glue("1", ".csv")))` returns `TRUE` if the file exists and `FALSE` if it doesn't)

```{r prob6}
iris %>%
  nest(data = -Species) %>%
  pmap(~write_csv(.y, path = here::here(glue::glue(.x, ".csv"))))
```

Your solution to add a unit test to the code above (using a `testthat` function):

```{r, soln6, error=TRUE}
# ADD SOLUTION HERE
# BEGIN SOLUTION
test_that('files should exist', {
  map(1:3, ~ expect_true(file.exists(here::here(glue::glue(.x, ".csv")))))
  })
# END SOLUTION
```