---
title: "Lecture Notes: Data Wrangling Part II"
subtitle: "Version 1.0.0"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```


## Learning Objectives

From today's class, students are anticipated to be able to:

+ Grouping data to compute summarized functions using `dplyr::group_by()` and `dplyr::summarize()`.
+ Compute new columns using `mutate()`.
+ Use `dplyr::across()` to use functions with specific columns.
+ Use scoped variants of `summarise()` and `mutate()`, with `across()`.
+ Apply your dplyr knowledge to exploratory data analysis of a dataset.

## Parallel Resources

This topic parallels the following resources. Take about 30 minutes to get acquainted with the following before coming to class:

- [STAT 545 Episode 6-A: dplyr Part 2: Calculations on tibbles](https://youtu.be/38BA_v7BwKk)
- We recommend you read / skim through the following resources for more details
   - [Jenny's "dplyr single table verbs"](https://stat545.com/dplyr-single.html) for dplyr verbs.
   - [Hadley's "across" blog post](https://www.tidyverse.org/blog/2020/04/dplyr-1-0-0-colwise/) to learn more about `across()`

We suggest referring to this material when engaging with today's topic.

## Additional Resources

- [dplyr vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html)
- [Jenny's dplyr tutorial](https://stat545.com/dplyr-single.html)

## Quiz

True or False?

1. `summarise()` reduces each group down to one row, and `mutate()` preserves the number of rows. But, between the two, there's a way to reduce each group down to _two_ rows.
2. `across()` is useful, but it doesn't allow you to make new column names. 


## Demonstration


```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(gapminder)
gapminder
```


### Demo 1

Compute a weighted mean of life expectancy for different years, weighting by population.

```{r}
gapminder %>% 
  group_by(year) %>% 
  mutate(wmean = sum(lifeExp * pop) / sum(pop))
```

What if I wanted those values repeated for each line?


### Demo 2

For each continent and year, how many countries have a "low" life expectancy? Let's say life expectancy is "low" if it's less than the median most recent life expectancy for each continent. 

```{r}
v <- c(4, 7, 2, 4, 5, 9)
v[v < 6]
v < 6
v[1:2]

gapminder %>% 
  group_by(continent) %>% 
  mutate(median = median(lifeExp[year == 2007])) %>% 
  filter(lifeExp < median) %>% 
  count(continent, year)
  # group_by(continent, year) %>% 
  # summarise(n = n())

gapminder %>% 
  mutate(asia = continent == "Asia")
```

### Demo 3

Get the maximum and minimum `gdpPercap` for all continents.

```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarise(max_g = max(gdpPercap),
            min_g = min(gdpPercap)) %>% 
  group_by(continent == "Asia") %>% 
  summarise(max2 = max(max_g))
```


### Demo 4

Filter gapminder to all entries that have experienced a drop in life expectancy. Hint: `dplyr::lag()`.

```{r}
v <- 1:4
v
dplyr::lag(v)

gapminder %>% 
  group_by(country) %>% 
  arrange(year) %>% 
  mutate(inc = lifeExp - lag(lifeExp)) %>% 
  filter(inc < 0)
```


### Demo 5

Filter gapminder to contain six rows: the rows with the three largest GDP per capita, and the rows with the three smallest GDP per capita. Don't make any excerpts.

```{r}
gap1 <- filter(gapminder, continent == "Europe")
gap2 <- summarise(gap1, mean_gdp = mean(gdpPercap))
gap2

nrow(gapminder)

gapminder %>% 
  arrange(gdpPercap) %>% 
  slice(c(1   , 2   , 3, 1702, 1703 ,1704)
    #c(1:3, (n() - 2):n())
  )

c(1:3, (1704 - 2):1704)
c(1:3, 1702:1704)


gapminder %>% 
  group_by(continent) %>% 
  mutate(n_function = n())
```


```{r}
gapminder %>% 
  mutate(decade = floor(year / 10)) %>% 
  group_by(country, continent, decade) %>% 
  summarise(across(where(is.numeric), mean)) %>% 
  select(-year)
```


