---
title: "Lecture Notes: Plotting Part II"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```




### Demo 1

From last class:

> Compute a weighted mean of life expectancy for different years, weighting by population.

Plot that global average life expectancy over time.

```{r}
gapminder %>% 
  group_by(year) %>% 
  summarise(wmean = sum(lifeExp * pop) / sum(pop)) %>% 
  ggplot(aes(year, wmean)) +
  geom_line()
```

### Demo 2

From last class:

> For each continent and year, how many countries have a "low" life expectancy? Let’s say life expectancy is "low" if it’s less than the median most recent life expectancy for each continent.

Plot the number of "low life expectancy" countries over time for each continent. 

```{r}
gapminder %>% 
  group_by(continent) %>% 
  mutate(median = median(lifeExp[year == 2007])) %>% 
  filter(lifeExp < median) %>% 
  count(continent, year) %>% 
  ggplot(aes(year, n)) +
  geom_line(aes(colour = continent)) +
  ylab("# of countries with low lifeExp") +
  labs(title = "Evolution over time") +
  scale_x_continuous("") +
  scale_colour_discrete("") +
  theme_minimal() +
  theme(panel.grid = element_blank())
```


### Demo 3

**(a)**

Plot GDP per capita over time, identifying each continent. 

```{r}
p3 <- ggplot(gapminder, aes(year, gdpPercap)) +
  geom_line(aes(group = country, colour = continent), alpha = 1/3) +
  facet_wrap(~ continent) +
  scale_y_log10(labels = scales::dollar_format()) +
  guides(colour = FALSE)
p3
```


**(b)**

From last class Demo 3:

> Get the maximum and minimum `gdpPercap` for all continents.

We'll also add _for each year_.  Today: modify plot (a) above so that the lines for each continent are contained within a band.

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  mutate(max = max(gdpPercap),
            min = min(gdpPercap)) %>% 
  ggplot(aes(year, (gdpPercap))) +
  geom_ribbon(aes(ymin = min, ymax = max), alpha = 0.5, fill = "gray") +
  geom_line(aes(group = country, colour = continent), alpha = 1/3) +
  # geom_ribbon(aes(ymin = quantile(gdpPercap, probs = 0.25),
  #                 ymax = quantile(gdpPercap, probs = 0.75),
  #                 group = year)) +
  facet_wrap( ~ continent) +
  scale_y_log10(labels = scales::dollar_format()) +
  guides(colour = FALSE) +
  theme_bw()
```


### Demo 4

From last class:

> Filter gapminder to all entries that have experienced a drop in life expectancy. Hint: `dplyr::lag()`.

Today: Was a drop in life expectancy just as likely for any year, or were some years more prone to see a drop than others? Is this true for all continents?

```{r}
gapminder %>% 
  group_by(country) %>% 
  arrange(year) %>% 
  mutate(inc = lifeExp - lag(lifeExp)) %>% 
  filter(inc < 0) %>% 
  ggplot(aes(year)) +
  geom_bar()
```


### Demo 5

How has the distribution of GDP per capita changed over time for each continent?

(Breakout rooms?)

```{r}
gapminder %>% 
  filter(continent != "Oceania") %>% 
  ggplot(aes(gdpPercap, factor(year))) +
  ggridges::geom_density_ridges(aes(fill = continent), alpha = 1/3) +
  scale_x_log10("GDP per capita", labels = scales::dollar_format()) +
  ylab("Year") +
  theme_minimal() +
  scale_fill_discrete("")
```


### Effective Vis

Let's go through [Principles of Figure Design](https://vincenzocoia.com/communicating_data/#28) slides, modelling Part II of [Claus Wilke's vis book](https://clauswilke.com/dataviz/).

### Demo 6

The following mock dataset marks the (x,y) position of a caribou at four time points. 

```
tribble(
  ~time, ~x, ~y,
  1, 0.3, 0.3,
  2, 0.8, 0.7,
  3, 0.5, 0.9,
  4, 0.4, 0.5
) %>% 
  ggplot(aes(x, y)) + 
  geom_line()
```

- Fix the plot below so that it shows the path of the caribou. 
- Add an arrow with `arrow = arrow()`.
- Add the `time` label with `geom_text()`.

```{r}
tribble(
  ~time, ~x, ~y,
  1, 0.3, 0.3,
  2, 0.8, 0.7,
  3, 0.5, 0.9,
  4, 0.4, 0.5
) %>% 
  ggplot(aes(x, y)) + 
  geom_path(arrow = arrow()) +
  geom_text(aes(label = time), nudge_x = 0.02) +
  theme_void()
```

### Demo 7

- Change the following plot so that it shows _proportion_ on the y-axis, not count.
- Change the x-axis so that it doesn't appear to be continuous.
- Put the bars for transmission side-by-side with their own colour.
- Capitalize the legend title.

```
mtcars %>% 
  mutate(transmission = if_else(am == 0, "automatic", "manual")) %>% 
  ggplot(aes(cyl)) +
  geom_bar(aes(colour = transmission))
```

```{r, fig.width = 5, fig.height = 2}
mtcars %>% 
  mutate(transmission = if_else(am == 0, "automatic", "manual")) %>% 
  ggplot(aes(factor(cyl))) +
  geom_bar(aes(y = ..prop.., fill = transmission, group = transmission), 
           position = "dodge") +
  scale_fill_discrete("Transmission") +
  labs(y = "Proportion", x = "Cylinders") +
  theme_minimal()
```

### Demo 8

Here's the number of people having a certain hair colour from a sample of 592 people:

```{r}
(hair <- as_tibble(HairEyeColor) %>% 
  count(Hair, wt = n))
```

Fix the following bar plot so that it shows these counts.

```
ggplot(hair, aes(Hair, n)) +
  geom_bar()
```

```{r}
ggplot(hair, aes(Hair, n)) +
  geom_col()
```

### Demo 9

Here's the number of people having a certain hair and eye colour from a sample of 592 people:

```{r}
(hair_eye <- as_tibble(HairEyeColor) %>% 
  count(Hair, Eye, wt = n))
```

### Demo 10
 
Fix the following plot so that it shows a filled-in square for each combination. 

```
ggplot(hair_eye, aes(Hair, Eye)) +
  geom_point(aes(colour = n))
```

```{r}
ggplot(hair_eye, aes(Hair, Eye)) +
  geom_bin2d(aes(fill = n)) +
  theme_minimal()
```

