---
title: "Reading and Writing Data"
output: html_document
---

From today's class, students are anticipated to be able to:

- Read and write a delimited file, like a csv, from R using the `readr` package.
- Make relative paths using the `here::here()` function.
- Read data from a spreadsheet
- Read and write R binary files (rds files) from R.


## Resources

Video lecture:

- [Reading and Writing Data](https://youtu.be/2nHuNu6Zaqk)

Tutorial:

- The ["Writing and Reading files"](https://stat545.com/import-export.html) chapter of stat545.com.

## Formats 

Here are a few formats that you might want to read and write data to: 

- *Spreadsheets*: Excel, Google Sheets, Numbers 
- *Delimited files*: Plaintext files containing data, e.g. csv, tsv
- *R binary*: A serialization of an R object to a binary file. Basically, that means that it can be loaded in and out of R, but it can't be opened by anything but R. 

Thumb rules: 

- csv are the most "one-size-fits-all": you can open them in spreadsheet software, but they are also plaintext, so are lightweight, can be opened in any text editor, and can be "diff"ed. 
- Spreadsheets are nice for human interaction. 
- R binary is VERY niche. Don't reach for it unless none of the other options suit your purposes. 


## Reading and writing data in R 

We use the `readr` package for this, because we think it has the most "work right out of the box" experience. 

Main functions of note are `read_csv()` and `write_csv()`: tidyverse equivalents of the base R functions `read.csv()` and `write.csv()`. 

Want to read and write to an Excel file? The `readxl` package in the tidyverse is for you! 

For the very niche option of R binary: `read_rds()` and `write_rds()`. 

## Test your Understanding

1. True or False: if you want to be deliberate about where `here::here()` points to on your computer, you need to ensure you have an .Rproj file.
2. True or False: Suppose you have an .Rproj file in the same folder as your R script. Running `here::here()` from that R script will always point to that folder.

---

## Your turn: try it out! 

Open RStudio. Go to Session => Set Working Directory => Choose Directory and then pick a folder you would like to read and write data into. Then, run the following piece of code: 

```{r, eval = FALSE} 
library(tidyverse) 
library(gapminder)

gap_asia_2007 <- gapminder %>% 
  filter(year == 2007, continent == "Asia")
head(gap_asia_2007)
```
Write `gap_asia_2007` to a comma-separated value (csv) file named `exported_file.csv` with just one command: 

```{r, eval = FALSE} 
write_csv(FILL_THIS_IN, "exported_file.csv")
```

Check out your files after executing this line! 

Now, let's practice reading csvs by reading the file we just wrote back into R: 

```{r, eval = FALSE} 
gap_asia_2007_in <- read_csv("FILL_THIS_IN")
```

Check out your R environment after executing this line!

Also notice the output of running `read_csv`. This tells us about the types of variables that were read in. It's a good habit to check this every time you run a `read_` function. Sometimes we might want to change how these variable types are specified...

## Main idea behind `here::here()`

We just wrote and read files to our current directory. If we wanted to use a different folder on our computer, we could specify something like: 

- `Documents/STAT545/exported_file.csv` - Mac uses forward slashes
- `Documents\STAT545\exported_file.csv` - Windows uses backward slashes

However, if you wanted to make your Rproj more portable and accessible to more users in a cross-platform (between Mac, Unix, Windows users), rather than specifying every path explicitly, `here::here()` allows you to set relative paths much more easily.


For a quick start, check out [this rant](https://github.com/jennybc/here_here) by [Jenny Bryan](https://jennybryan.org/), the founder of STAT 545,

For a longer written tutorial, read the above noted STAT 545 chapter on ["Writing and Reading files"](https://stat545.com/import-export.html). 