---
title: "Lecture Notes: Tidy Data"
subtitle: "Version 1.0.0"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```


## Learning Objectives

From today's class, students are anticipated to be able to:

- recognize whether a given dataset is 'tidy' or 'untidy'
- convert a dataset between the 'long' and 'wide' format, using `tidyr::pivot_longer()` and `tidyr::pivot_wider()`
- assess which format is best suited for each type of analysis
- deal with missing data in a tibble


## Parallel Resources

This topic parallels the following resources. Take about 30 minutes to get acquainted with the following before coming to class:

- [STAT 545 Episode 8-A: tidyr for Pivoting and Tidy Data](https://youtu.be/qivE6exNsZI) (also available on canvas)
- Read the [`pivot` vignette](https://tidyr.tidyverse.org/articles/pivot.html) with `tidyr`

We suggest referring to this material when engaging with today's topic.

## Additional Resources

- `tidyr` [vignette](https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html) on tidy data for more depth on tidy data.
- [Hadley's Paper on Tidy Data](https://vita.had.co.nz/papers/tidy-data.pdf) for a formal journal article on tidy data.
- Other new [additions](https://www.tidyverse.org/articles/2019/09/tidyr-1-0-0/) in `tidyr` v1.0.0


## Quiz

1. True or False: it's possible to lengthen a tibble "all the way" so that it can't be lengthened anymore.
2. True or False: To make a scatterplot viewing the relationship of Africa's GDP per capita vs. Europe's, use `pivot_wider()`. 

```{r}
dat <- tibble(x = letters)
dat %>% 
  pivot_longer(cols = x) %>% 
  pivot_longer(cols = everything())

gapminder %>% 
  group_by(continent) %>% 
  summarise(gdpPercap = mean(gdpPercap))  %>% 
  pivot_wider(names_from = continent, values_from = gdpPercap)
```


## Demonstration


```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(palmerpenguins)
library(gapminder)
knitr::opts_chunk$set(error = TRUE)
guestlist <- read_csv(here::here("content", "data", "wedding", "attend.csv"))
```

### Review

#### Collaborative Milestone 2

Logical operators: `%in%`, `!=`, `!`

```{r}
v <- c("foo", "foo", "bar", "baz")
v %in% c("foo", "baz")
v == "foo" | v == "baz"
v == "foo"
v == c("foo", "baz")
v == c("foo", "baz", "foo", "baz")
v != "foo"
!(v == "foo")
```


- Avoid using deprecated functions: use `across()` instead of scoped variants like `mutate_at()` and `summarise_if()`.
- Troubleshooting? Check the function's documentation.
  - Easy to access with `?function`
  - Typically concise. Not like those old fashioned "User Manuals" [like this WinBUGS manual](https://www.mrc-bsu.cam.ac.uk/wp-content/uploads/manual14.pdf).

#### Worksheet 3-A

Any requests? If not, Question 1.3:

> In a single expression, make a tibble with the following columns for each island in the penguins data set:
> 
> - What is the mean of each numeric variable in the penguins dataset in each island? Keep the column names the same.
> - How many penguins are there in each island? Add this to a column named `n`.

```{r}

```

#### 1.5

```{r}
mean(c(1, 4, 3, NA), na.rm = TRUE)

penguins %>% 
  mutate(across(where(is.numeric), mean, na.rm = TRUE, trim = 0.2))

penguins %>% 
  mutate(across(where(is.numeric), ~ mean(.x, na.rm = TRUE, trim = 0.2)))

(answer1.5 <- penguins %>%
 mutate(across(where(is.numeric), ~ if_else(is.na(.x), mean(.x, na.rm = T), as.numeric(.x)))))
```


### tidyr Demo

Let's work with this guestlist dataset.

```{r}
guestlist
```

#### 1. Country relationship

Using the `gapminder` dataset: What's the relationship between Canada's GDP per capita and the United Kingdom's? First, identify the tidy format.

```{r}
gapminder %>% 
  filter(country %in% c("Canada", "United Kingdom")) %>% 
  pivot_wider(id_cols = year, names_from = country, values_from = gdpPercap) %>% 
  ggplot(aes(x = Canada, y = `United Kingdom`)) +
  geom_point()
gapminder %>% 
  filter(country %in% c("Canada", "United Kingdom")) %>% 
  ggplot(aes(x = gdpPercap, y = gdpPercap)) +
  geom_point()

`_x` <- 4
`this is a variable` <- 5
```

#### 2. Who's PENDING?

Our task: extract all "PENDING" entries, paired with who it is, and what information is pending.

What is a tidy data format for this question? Do the tidying.

```{r}
guestlist %>% 
  pivot_longer(-c(party, name), names_to = "info", values_to = "status") %>% 
  filter(status == "PENDING")
```


#### 3. Relationship between PENDING

What's the relationship between whether or not a meal is PENDING, and the attendance status? First determine what a tidy format is for this question, then answer the question. 

```{r}
guestlist
```


```{r}
guestlist %>% 
  pivot_longer(-c(party, name), names_to = c(".value", "event"), names_sep = "_")
```

#### 4. Split name

Split the guest name into first and last name.

```{r}
guestlist %>% 
  separate(col = name, into = c("first", "last"), sep = " ")
```




### Deprecated functions: spread and gather

Pivoting used to be achieved through the tidyr `spread()` and `gather()` functions. These are deprecated and should not be used. 
