---
title: "List Columns"
output: html_document
---

Did you know that a column of a tibble can contain a list -- not just a vector? This week explores when such a thing would come in handy, and how to work with list columns. 

From this topic, students are anticipated to be able to:

- Use the `map` family of functions from the purrr package to iteratively apply a function.
- Create and operate on list columns in a tibble using `nest()`, `unnest()`, and the `map()` family of functions.
- Define functions on-the-fly within a `map` function using shortcuts.
- Apply list columns to cases in data analysis: columns of models, columns of nested lists (JSON-style data), and operating on entire groups within a tibble. 

## Resources

Video lectures:

- [Vectors and Lists in R](https://www.youtube.com/watch?v=nJgrDoCENBI)
- [R List Columns: purrr map and nesting](https://www.youtube.com/watch?v=_731PZ8I9m0)

Written material:

- [R4DS Chapter 21: Iteration](https://r4ds.had.co.nz/iteration.html) for purrr
    - 21.1 for an intro
    - 21.5 for the `map` family of functions
    - The intro of 21.7 for the `map2` and `pmap` families.
- ["List Columns"](https://jennybc.github.io/purrr-tutorial/ls13_list-columns.html) from [Jenny's purrr tutorial](https://jennybc.github.io/purrr-tutorial/index.html)
- ["Nested data" article](https://tidyr.tidyverse.org/articles/nest.html) on tidyr's website.

Want to dig deeper? These resources can help.

- [Advanced R Chapter 9: Functionals](https://adv-r.hadley.nz/functionals.html) -- looking at purrr and `map()` from a programming perspective. 
- [tidyr's rectangling vignette](https://tidyr.tidyverse.org/articles/rectangle.html) -- for handling deeply nested lists (JSON-style data), similar to tidyr's `pivot_` functions. 

## Test Your Understanding

1. True or False: `map(1:3, ~ function(x) x ^ 2)` returns the list `list(1, 4, 9)`.
2. You write a function `square()` that squares its input -- but the first thing it does is print a message to the screen! True or False: `map_dbl(1:10, square)` will throw an error, because the output is not a `dbl` (a number) -- it contains the message, too.
3. True or False: purrr-style functions, starting with `~`, can be used in dplyr's `across()` function, such as `mutate(across(where(is.numeric), ~ .x ^ 2))`, and this function can only take one argument (`.x`)
4. True or False: If I have 10 tibbles I want to save to file, and they're all stored in a list, the best purrr function to use for saving these to file is `map()`.
5. True or False: Just as `c(c(1, 2), c(3, 4))` returns the vector of numbers from 1 to 4, `list(list(1, 2), list(3, 4))` returns the list of numbers from 1 to 4. 
6. True or False: `tibble(model = lm(mpg ~ wt, data = mtcars))` doesn't work because it doesn't use a `map` function.

## Demonstration

```{r}
library(tidyverse)
library(palmerpenguins)
library(distplyr)
library(repurrrsive)
```

We will practice by jumping right into the worksheet. Here are some additional exercises that we may cover in class.

By the way, why use the `map` functions over `for` loops? Aside from what's said at [R4DS's Section 21.4](https://r4ds.had.co.nz/iteration.html#for-loops-vs.-functionals), the `map` family encourages the input vector(s) and function to be carefully defined first, whereas a `for` loop is more of a free-for-all. Using `map()` is also more readable, is pipe-friendly, and jibes well with list columns in tibbles. 

### Demo 1: `map2`

Using the `palmerpenguins::penguins` tibble: 

1. Make a Normal distribution using `distplyr::dst_norm()` for the body mass of each species, using estimates taken from `mean()` and `var()`. Title the column `distribution`.
2. Calculate the 0.975-quantile from each distribution using `distplyr::eval_quantile()`. Try this by making a new column of distributions, and by not.

Starter code:

```
(demo1 <- penguins %>% 
  group_by(species) %>% 
  summarise(mean = mean(body_mass_g, na.rm = TRUE),
            var  = var(body_mass_g, na.rm = TRUE)))
```

Answer:

```{r}
(demo1a <- penguins %>% 
  group_by(species) %>% 
  summarise(mean = mean(body_mass_g, na.rm = TRUE),
            var  = var(body_mass_g, na.rm = TRUE)) %>% 
   mutate(distribution = map2(mean, var, dst_norm),
          quantile = map_dbl(distribution, eval_quantile, at = 0.975)))

(demo1b <- penguins %>% 
  group_by(species) %>% 
  summarise(mean = mean(body_mass_g, na.rm = TRUE),
            var  = var(body_mass_g, na.rm = TRUE)) %>% 
    mutate(quantile = map2(mean, var, dst_norm) %>% 
             map_dbl(eval_quantile, at = 0.975)))
```

Explore distplyr more:

```{r, fig.width = 3, fig.height = 3}
chin <- demo1a$distribution[[2]]
print(chin)
plot(chin, "density")

eval_density(chin, at = 3000:3010)
eval_quantile(chin, at = c(0.25, 0.5, 0.75))
enframe_quantile(chin, at = c(0.25, 0.5, 0.75))

(x <- c(chinstrap = 3, gentoo = 7, adelie = 5))
enframe(x)
```


### Demo 2: `unnest()`

`unnest()` need not always be paired with `nest()`. For the above distributions, evaluate the 0.25, 0.50, and 0.75 quantiles using the `distplyr::enframe_quantile()` function.

Starter code:

```
demo1a %>% 
  mutate(quantile = FILL_THIS_IN)
```

Answer

```{r}
demo1a %>% 
  mutate(quantile = map(distribution, enframe_quantile, at = c(0.25, 0.5, 0.75))) %>% 
  unnest(quantile)
```

Nesting the penguins tibble

```{r}
penguins %>% 
  nest(data = body_mass_g)
```

```{r}
penguins %>% 
  nest(data = body_mass_g) %>% 
  mutate(s = map_dbl(data, ~ sqrt(.x$body_mass_g)))
penguins %>% 
  mutate(across(where(is.numeric), ~ sqrt(.x - 10)))
```



### Demo 3: `i` for accessing index

*Time permitting*

Sometimes the vector/list we're iterating over has names, and it's useful to use those names. To access these names, use the `imap` family.

Below is a named list of tibbles. Save them to file using the appropriate purrr function, using the names as the file names.

```{r}
data <- list(mtcars   = mtcars, 
             chickwts = chickwts, 
             pressure = pressure)
iwalk(data, ~ print(.y))
#iwalk(data, ~ write_csv(.x, glue::glue(.y, ".csv")))
```

### Demo 4: `unnest_wider()` and `unnest_longer()`

*Time permitting*

1. Explore the `repurrrsive::got_chars` nested list. It contains information about Game of Thrones characters.
2. Put the list in a tibble.
3. Would widening the list column work best, or lengthening? Do it.

Explore:

```{r}
str(got_chars, list.len = 4)
```
Make tibble:

```{r}
(dat <- tibble(character = got_chars))
```

`unnest_wider()` is what we want, not `unnest_longer()`:

```{r}
dat %>% 
  unnest_wider(character)
dat %>% 
  unnest_longer(character)
```

