---
title: "Special objects in R: Dates, Times, and Factors"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```

```{r, echo=FALSE, message=FALSE} 
library(tidyverse)
library(palmerpenguins)
```

From today's class, students are anticipated to be able to:

- Reorder levels within factors according to various principles.
- Make a date and datetime object and extract components from these objects, using the lubridate package.

## Resources

Video lecture:

- [Special Data Types in R: Dates, Times, and Factors](https://youtu.be/TTNfLtEUFDQ)

Other resources, in addition to the notes below:

- For factors:
    - The [stat545.com](https://stat545.com/factors-boss.html) chapter on Factors
    - The [forcats](https://forcats.tidyverse.org) package page and reference guide on page. 
- For dates and times:
    - The [R4DS](https://r4ds.had.co.nz/dates-and-times.html) chapter on Dates and Times
    - The `tsibble` [vignette](https://cran.rstudio.com/web/packages/tsibble/vignettes/intro-tsibble.html) to learn more about embedding a time series within a tibble. 

## Dates and/or Times 
### Date/Time Objects in R 

#### Exercise 

- Write down today's date. (Don't peek at your neighbours!)

- Copy what you wrote down into the Google Sheets link I posted to the #lectures channel on Slack. 

We'll look at the results together - I think they will speak for themselves ... 

#### Code Demo

Let's say you want to convert a string into a date object. 

The `lubridate` package in the `tidyverse` has your back with a family of functions named as permutations of "y", "m", and "d". 

```{r, message=FALSE} 
library(lubridate)
ymd("2020-10-06") 
mdy("October 6th, 2020")
dmy("06-10-2020")
```

We can also convert strings into date-time objects using `lubridate` functions. The family of functions you're looking for is named with a "y"/"m"/"d" permutation followed by an "_" and one or more of "h", "m", and "s". 

```{r} 
ymd_hms("2020-10-06 14:20:39")
ymd_hm("2020-10-06 14:20")
```

Note that there are also time objects, though we don't cover them here. 

Once you've acquired a date or date/time object, you can use `lubridate` functions to extract individual components of the data or date/time. 

```{r} 
today_date <- today()
year(today_date)
wday(today_date, label=TRUE)
yday(today_date)
```

### Date/Time Math 

An incredible advantage of working with date and date-time objects in R is that you can make R do pesky time arithmetic for you. 

```{r} 
today_date <- today()
christmas_this_year <- mdy(paste("December 25", year(today_date)))

christmas_this_year - today_date
```

If you want to make sure your difference comes out as seconds (or want to convert between different units) you can use `as.duration`: 

```{r}
as.duration(christmas_this_year - today_date)
```


## Factors

> "There is no other object that creates as much trouble as factors." - Patrick Burns, "The R Inferno". 

In R, we use *factors* to represent *categorical variables*: variables that take on a fixed number of known values (i.e. *levels*). For example, in the `penguins` data set, `species` is a factor with three levels: "Adelie", "Chinstrap", and "Gentoo".

```{r} 
glimpse(penguins)
```

**Under the hood**, R stores a factor with (say) 3 levels as a numeric vector containing integers between 1 and 3, paired with a character vector of length 3 that identifies the mapping between the numbers 1, 2, and 3 and the levels. 

This is not immediately obvious, because R will print factors using the character string levels rather than the numbers that it stores: 

```{r} 
penguins %>% slice_sample(n=10) %>% pull(species)
```

This dual nature of factors creates a whole slew of hidden traps and headaches, especially for new useRs! 

Nevertheless, factors are important and worth the pain. Many functions throughout the R landscape expect categorical variables to be coded as factors. For example, when making plots in either `ggplot2` or in base R, we need factors in order to map categorical variables to aesthetic elements like colour. 

To make our lives easier, we will work with factors through the `forcats` package loaded as part of the `tidyverse`.  

### Demonstration 

We will work on Worksheet A5 (factors and dates portion).

## Test Your Understanding

1. The output of the following code is equivalent to a factor with the letters "a" to "f".

``` r
(abc <- factor(letters[1:3]))
#> [1] a b c
#> Levels: a b c
(def <- factor(letters[4:6]))
#> [1] d e f
#> Levels: d e f
c(abc, def)
```

2. The output of the following code is a date object.

```r

library(lubridate)

date <- ymd("2020-10-13")

dt <- ymd_hms("2020-10-13 09:30:00")

c(date, dt)

```
