---
title: "R Functions for Data Analysis"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```

From this topic, students are expected to be able to:

- Start getting a sense of when to make a function in a data analysis (we will build on this next week).
- Workflow for building a function: start interactively, wrap it as a function. `return()`. Argument names. 
- Fortify a function: 
  - generalize the function and use smart defaults; `NA` handling, and ellipses package [https://ellipsis.r-lib.org/](https://ellipsis.r-lib.org/)
  - provide useful error messages; sidebar: `if` statements
  - Unit tests, and (sidebar) assertions
- Data masking in a function.
- Documenting a function

Possible topics this week, time permitting:

- `try()` and `withTimeout()`.
- environments and scoping. Perhaps `search()`; `with()`.

What we're _not_ covering (this week):

- Documenting functions with roxygen2 (R packages)
- S3 objects (and other object oriented styles) (R packages)


## Resources

Video lecture: 

- [R Functions for Data Analysis](https://youtu.be/3nDgR7l5Tps)

Written resources:

- [stat545.com Functions, Parts 1-3](https://stat545.com/functions-part1.html)
- [R4DS functions chapter](https://r-pkgs.org/setup.html#setup-tools)

## Test Your Understanding

Consider the following code:

```
x <- 1
f <- function(t) {
  x <- -1
  if (x > 0) return(t)
  x * t + 1
  x * t
}
f(1)
```

1. True or False: the output of `f(1)` is `-1`. 
2. True or False: after running the above code, `x` is `-1`. 
3. True or False: Curly-curly `{{}}` can be used whenever we want to refer to column names outright, such as when subsetting in base R with `[[`.
4. You've made a function `quick_scatter()` that you can pipe a tibble into, and outputs a scatterplot of two variables. So far, you're only using the function to explore the data for your own eyes (perhaps even putting your code directly in the console). True or False: at this point, you should stop and make sure your function is documented and has formal unit tests. 


## Functions: an example

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(gapminder)
library(testthat)
```

Sometimes it's useful to work with the _log_ of a variable with a positive skew, such as `gapminder`'s GDP per capita:

```{r}
cowplot::plot_grid(
  ggplot(gapminder, aes(gdpPercap)) +
    geom_histogram() +
    theme_minimal(),
  ggplot(gapminder, aes(log(gdpPercap))) +
    geom_histogram() +
    theme_minimal()
)
```

And, it's sometimes useful to scale a variable to have mean 0 and variance 1. We'll do both:

```{r}
gapminder %>% 
  mutate(gdpPercap_scaled = (log(gdpPercap) - mean(log(gdpPercap))) / sd(log(gdpPercap)),
         pop_scaled = (log(pop) - mean(log(pop))) / sd(log(pop)))

gapminder %>% 
  mutate(gdpPercap = (log(gdpPercap) - mean(log(gdpPercap))) / 
           sd(log(gdpPercap)),
         pop = (log(pop) - mean(log(pop))) / sd(log(pop)))
```

Where might there be repetition in this analysis? Where __is__ there repetition?

- Doing this scaling over and over again throughout the analysis.
- Repetition across multiple variables, like `pop` and `lifeExp`
- `gdpPercap` appears three times.

Solve the problem of repetition by writing _functions_.

Turn the above into a function.

```{r}
my_scale <- function(x, remove_na = TRUE, verbose = FALSE, ...) {
  if (verbose) cat("Taking the log...\n")
  logx <- log(x)
  if (verbose) cat("Done taking the log!\n")
  (logx - mean(logx, na.rm = remove_na, ...)) / sd(logx, na.rm = remove_na)
}
my_scale(1:10)
```

Now we can use the function in our previous dplyr workflow:

```{r}
gapminder %>% 
  mutate(gdpPercap = my_scale(gdpPercap),
         pop = my_scale(pop))
```

Each function you create shouldn't do too much, but rather should do one thing well. Don't be like me during my PhD and [make a gigantic `cnqr()` function](https://github.com/vincenzocoia/cnqr/blob/master/R/cnqr.R) that tries to do everything. Modularizing computation into smaller functions makes troubleshooting much easier.

## Testing your functions

`testthat` is a collection of functions developed by Hadley Wickham that makes unit testing easy for developers. You can read more about the structure of tests [here](https://r-pkgs.org/tests.html#test-structure). In a nutshell, tests are organized hierarchically: **expectations** are grouped into **tests**. Functions that start with `expect_` describe the expected result of a computation (e.g. Does it have the right class?) - these are expectations. Tests are created with `test_that()` and they group together multiple *expectations* to test the output of a function (at their simplest level). 

Let's test the `my_scale()` function we made earlier, first informally.

```{r}
my_scale(exp(c(1, 3)), verbose = TRUE)
my_scale(exp(c(1, 3, NA)))
my_scale(1:10)
my_scale(1:10, trim = 0.1, foobar = 3)
```

Now, write a formal test that doesn't require eyeballs on the code:

```{r}
test_that("Basic testing", {
  expect_equal(my_scale(exp(c(1, 3))), (c(1, 3) - 2) / sqrt(2))
  expect_equal(my_scale(exp(c(1, 3, NA))), (c(1, 3, NA) - 2) / sqrt(2))
  expect_error(my_scale(letters))
})
```

Testing your functions is a way of recognizing that your data analysis will evolve, and is a way of knowing which of your functions need updating if you make a breaking change in your analysis. 

## A return to dplyr scoped variants with `across()`

Let's try our function applied to multiple columns of the `gapminder` dataset using dplyr's `across()` function. 

```{r}
gapminder %>% 
  mutate(across(c("gdpPercap", "pop"), my_scale))
```

## Function arguments without quotation marks: curly-curly

Sometimes your function needs to take variable names without quotation marks. For example, `select(gapminder, lifeExp)` does not put quotation marks around `lifeExp` (the reason being that `lifeExp` is like a variable in our workspace, if we were to include column names in our R Environment).

Making your own functions having this capability involves a workaround. For example, consider making a function to produce a quick scatterplot between two columns in a dataset:

```{r}
quick_scatter <- function(data, x, y) {
  ggplot(data, aes(x, y)) +
    geom_point()
}
```

The function won't work:

```{r, error = TRUE}
quick_scatter(gapminder, x = gdpPercap, y = lifeExp)
```

The reason is that R is looking for variables named `gdpPercap` and `lifeExp` in the workspace, and cannot find them. To fix the problem, we can change the function definition so that `x` and `y` are embraced within two curly brackets -- an operator called "curly curly": 

```{r}
quick_scatter <- function(data, x, y) {
  ggplot(data, aes({{ x }}, {{ y }})) +
    geom_point()
}
```

Now the function will work:

```{r, fig.width = 3, fig.height = 2}
quick_scatter(gapminder, x = gdpPercap, y = lifeExp)
```

But, you can only use curly-curly when passing your function's argument to another function _that's anticipating a variable name without brackets_. For example, consider subsetting a data frame by a column name using `[[`, where quotation marks are needed:

```{r}
mtcars[["wt"]]
```

Here's a function that subsets a data frame using `[[`, not using curly-curly:

```{r}
my_subset <- function(data, variable) {
    data[[variable]]
}
```

The function is anticipating quotation marks, and will fail without them:

```{r, error = TRUE}
my_subset(mtcars, "wt") 
my_subset(mtcars, wt)
```

Using curly-curly doesn't grant the `[[` subsetting any data masking powers -- data masking must already be in place, as they are in `dplyr` and `ggplot2` functions:

```{r}
my_subset <- function(data, variable) {
    data[[{{ variable }}]]
}
```

We get an error if we don't include quotation marks:

```{r, error = TRUE}
my_subset(mtcars, wt)
```


