---
title: "List Columns"
output: html_document
---

```{r, message=FALSE}
library(tidyverse)
```


### Learning Objectives 

From this topic, students are anticipated to be able to:

- Use the `map` family of functions from the purrr package to iteratively apply a function.
- Create and operate on list columns in a tibble using `nest()`, `unnest()`, and the `map()` family of functions.
- Define functions on-the-fly within a `map` function using shortcuts.
- Apply list columns to cases in data analysis: columns of models, columns of nested lists (JSON-style data), and operating on entire groups within a tibble. 



## Resources

Video lectures:

- [Vectors and Lists in R](https://www.youtube.com/watch?v=nJgrDoCENBI)
- [R List Columns: purrr map and nesting](https://www.youtube.com/watch?v=_731PZ8I9m0)

Written material:

- [R4DS Chapter 21: Iteration](https://r4ds.hadley.nz/iteration.html) for purrr
    - 21.1 for an intro
    - 21.5 for the `map` family of functions
    - The intro of 21.7 for the `map2` and `pmap` families.
- ["List Columns"](https://jennybc.github.io/purrr-tutorial/ls13_list-columns.html) from [Jenny's purrr tutorial](https://jennybc.github.io/purrr-tutorial/index.html)
- ["Nested data" article](https://tidyr.tidyverse.org/articles/nest.html) on tidyr's website.

Want to dig deeper? These resources can help.

- [Advanced R Chapter 9: Functionals](https://adv-r.hadley.nz/functionals.html) -- looking at purrr and `map()` from a programming perspective. 
- [tidyr's rectangling vignette](https://tidyr.tidyverse.org/articles/rectangle.html) -- for handling deeply nested lists (JSON-style data), similar to tidyr's `pivot_` functions. 

## Test Your Understanding

1. True or False: `map(1:3, ~ function(x) x ^ 2)` returns the list `list(1, 4, 9)`.
2. You write a function `square()` that squares its input -- but the first thing it does is print a message to the screen! True or False: `map_dbl(1:10, square)` will throw an error, because the output is not a `dbl` (a number) -- it contains the message, too.
3. True or False: purrr-style functions, starting with `~`, can be used in dplyr's `across()` function, such as `mutate(across(where(is.numeric), ~ .x ^ 2))`, and this function can only take one argument (`.x`)
4. True or False: If I have 10 tibbles I want to save to file, and they're all stored in a list, the best purrr function to use for saving these to file is `map()`.
5. True or False: Just as `c(c(1, 2), c(3, 4))` returns the vector of numbers from 1 to 4, `list(list(1, 2), list(3, 4))` returns the list of numbers from 1 to 4. 
6. True or False: `tibble(model = lm(mpg ~ wt, data = mtcars))` doesn't work because it doesn't use a `map` function.

## Vectors vs Lists 

Here is a list in R; it holds multiple items. 

```{r} 
sample_list <- list(1:3, c("a", "b", "c"))
sample_list
```

A list might sound like a *vector*, which we have worked with before -- remember, we construct them using the `c()` function. Indeed, vectors and lists can both hold multiple items. But there are key differences. 

| Vectors | Lists |
|----------|------|
| Access elements with square brackets `[]` | Access elements with  `[[]]` | 
| Each element must be an atomic data type (i.e. a single value) | Elements can be anything, even another list or another vector | 
| Each element has to be of the same type | Elements can be as different as you like | 

## The Secret Life of Tibbles 

Did you know that data frames (and tibbles) are actually a special type of list? It's true! 

```{r} 
typeof(mtcars) 
typeof(palmerpenguins::penguins)
```

It turns out that they are actually *lists*, where each element of the list stores a column, which is either a list with the same number of entries as the tibble has rows, or a vector with the same number of entries as the tibble has rows. 

This has an important implication: we can efficiently apply a function to each column of a tibble by learning how to apply a function to each entry of a list. This is yet another way (beyond functions themselves) of avoiding duplicating code, which you will recall (from the functions topic) has many advantages. 

## Iteration

If you programmed before, you probably have an idea of how to do this with a `for` loop. Here's an example of a for loop in R that iterates over the entries of a numeric vector `x`, squares each entry, and stores the result in a numeric vector `output`: 

```{r} 
x <- 1:10 
output <- vector("double", length(x))

for(i in seq_along(x)) { 
    output[i] <- x[i]^2  
}

output

```

Often, you can replace loops with a compact call to a function in the `purrr` package. This has the advantage of making our code even more readable and compact, since we're expressing the same logic with less space. Here's an example using `purrr::map_dbl()` and a custom function: 

```{r} 
purrr::map_dbl(1:10, function(x) x^2)
```

The first argument specifies the list/vector we want to iterate over, and the second argument specifies a function that we want to apply to each entry. Options for specifying functions include the name of a function, a fully specified custom function (as demonstrated above), or one of the "shortcuts" the `purrr` developers have provided. 

Here are two examples of "shortcuts": 
```{r}
purrr::map_dbl(1:10, ~ (.x)^2)
purrr::map_dbl(1:10, \(x) x^2)
```

The second one is (IMO) easier to remember and appears to be the one that `purrr` developers are recommending now; see the [purrr cheatsheat](https://github.com/rstudio/cheatsheets/blob/main/purrr.pdf). But this change in recommendation appears to have happened around 2022/2023, so you may still see the first type of shortcut in many places in the wild. 

**Trivia**: [the story behind the name `purrr`](https://github.com/tidyverse/purrr/issues/35#issuecomment-71395551)

## Your turn: Worksheet B3, Part 1

We think the best way to get your bearings with `purrr` is to jump into Worksheet B3. Class 1 will be dedicated to getting your questions about Part 1 and about any concepts up to this point answered. 

## List Columns 

Did you know columns in a tibble can have type "list"? We call these types of columns "list columns".

Consider the following example: a snippet of the Game of Thrones data from [An API of Ice and Fire](https://anapioficeandfire.com/).  

```{r, echo = FALSE} 
got_lc <- unnest_wider(tibble(character = repurrrsive::got_chars), character) %>% select("name", "gender", "titles")
head(got_lc)
```

Some characters have one title (e.g. Will); others have more than one title (e.g. Theon Greyjoy). Consequently, the `titles` column is a *list column*, where each entry is a *list* that contains as many or as few strings as we like. 

## Your turn: Worksheet B3, Parts 2 and 3 

We think the best way to learn how to make and work with list columns (and get a taste for where they can be really useful!) is to jump back into the worksheet. 

Class 2 will be dedicated to getting your questions about Parts 2 and 3 and about any concepts involving list columns and nested lists answered. 


