---
title: "List Columns"
output: html_document
---

Did you know that a column of a tibble can contain a list -- not just a vector? This topic explores when such a thing would come in handy, and how to work with list columns. 

From this topic, students are anticipated to be able to:

- Use the `map` family of functions from the purrr package to iteratively apply a function.
- Create and operate on list columns in a tibble using `nest()`, `unnest()`, and the `map()` family of functions.
- Define functions on-the-fly within a `map` function using shortcuts.
- Apply list columns to cases in data analysis: columns of models, columns of nested lists (JSON-style data), and operating on entire groups within a tibble. 

## Resources

Video lectures:

- [Vectors and Lists in R](https://www.youtube.com/watch?v=nJgrDoCENBI)
- [R List Columns: purrr map and nesting](https://www.youtube.com/watch?v=_731PZ8I9m0)

Written material:

- [R4DS Chapter 21: Iteration](https://r4ds.had.co.nz/iteration.html) for purrr
    - 21.1 for an intro
    - 21.5 for the `map` family of functions
    - The intro of 21.7 for the `map2` and `pmap` families.
- ["List Columns"](https://jennybc.github.io/purrr-tutorial/ls13_list-columns.html) from [Jenny's purrr tutorial](https://jennybc.github.io/purrr-tutorial/index.html)
- ["Nested data" article](https://tidyr.tidyverse.org/articles/nest.html) on tidyr's website.

Want to dig deeper? These resources can help.

- [Advanced R Chapter 9: Functionals](https://adv-r.hadley.nz/functionals.html) -- looking at purrr and `map()` from a programming perspective. 
- [tidyr's rectangling vignette](https://tidyr.tidyverse.org/articles/rectangle.html) -- for handling deeply nested lists (JSON-style data), similar to tidyr's `pivot_` functions. 


## Vectors and Lists and Data Frames/Tibbles

In R, *vectors* hold multiple items/elements of the same atomic type: 

```{r} 
(v1 <- c(1, 5, 10)) # this is a vector holding items of type "double" 
(v2 <- c("abc", "def", "ghi")) # this is a vector holding items of type "character"
```

Here, atomic means "smallest unit":  a single integer or double, a single character string, a single true/false statement, etc. 

We access elements of a vector with square brackets: 
```{r}
v1[1]
v2[2]
```

We can find out how many elements are in a vector with the `length()` function: 
```{r} 
length(v1)
```

In R, *lists* also hold multiple items/elements, but each element can be *any* type. In fact, the elements don't even have to be atomic! 

```{r} 
(l1 <- list(1, TRUE))
(l2 <- list(palmerpenguins::penguins, 1:3, list(1:3, "blah")))
```

We access elements of a list with double square brackets: 
```{r} 
l1[[2]]
l2[[3]]
```

We can also find out how many elements are in a list with the `length()` function: 
```{r} 
length(l2)
```

As it turns out, data frames are actually a special type of list. This is true for tibbles too, since they're simply slightly fancier versions of data frames.

```{r} 
typeof(mtcars)
typeof(palmerpenguins::penguins)
```

This is why we can access columns using double square brackets - the data frame/tibble is actually a *list*, where each element is required to have the same length. 

```{r} 
mtcars[[1]]
palmerpenguins::penguins[["year"]]

length(mtcars[[1]])
length(mtcars[[2]])
```

## For loops and functionals

You may be familiar with `for` loops from prior experience with base R or with other programming languages. Here's an example of a for loop in R that iterates over the entries of a numeric vector `x`, squares each entry, and stores the result in a numeric vector `output`: 

```{r} 
x <- 1:10 
output <- vector("double", length(x))

for(i in seq_along(x)) { 
    output[i] <- x[i]^2  
}

output

```

For loops can often be replaced with a single call to a *functional* - a function that takes a function as an argument. To illustrate what I'm talking about, let's rewrite our example by defining a custom functional. 

```{r} 
square <- function(x) x^2 

my_map <- function(x, fun) { 
  output <- vector("double", length(x))
  for(i in seq_along(x)) { 
      output[i] <- fun(x[i])
  }
  
  output
}

my_map(1:10, square)
```

## `purrr` fundamentals demonstration

There's no need to write your own functionals for this purpose though - you can use the `map` family of functionals in the `purrr` package! 

```{r} 
library(purrr)
map_dbl(1:10, square)
```

Why use the `map` functions over `for` loops? The `map` family encourages the input vector(s) and function to be carefully defined first, whereas a `for` loop is more of a free-for-all. Using `map()` is also more readable, is pipe-friendly, and jibes well with list columns in tibbles. 

Fun trivia: [the story behind the name `purrr`](https://github.com/tidyverse/purrr/issues/35#issuecomment-71395551)

Let's jump into the worksheet to learn the basics of how to use `map` functions.

## List columns 

We now know that data frames and tibbles are just lists, where each element/column has the same length. We also know that an element of a list need not be an atomic - it could be an atomic vector or even a list! 

This means that an "entry" of a rectangular data frame or tibble could be *multiple* atomics, like so: 

```{r} 
list_column <- list(1:2, 3:4, 5:6)
regular_column <- c("a", "b", "c")

(my_df <- data.frame(col1 = I(list_column), col2=regular_column))

my_df[["col1"]]
my_df[1, 1]

library(tibble)
(my_tbl <- tibble(col1=list_column, col2=regular_column))
```

Base R data frames don't make it especially easy to make or work with list columns, so if you are planning to use list columns then tibbles are strongly recommended over data frames.  

Let's explore when list columns might be useful and how to manipulate list columns by jumping back into the worksheet. 

## Test Your Understanding

1. True or False: `map(1:3, ~ function(x) x ^ 2)` returns the list `list(1, 4, 9)`.
2. You write a function `square()` that squares its input -- but the first thing it does is print a message to the screen! True or False: `map_dbl(1:10, square)` will throw an error, because the output is not a `dbl` (a number) -- it contains the message, too.
3. True or False: purrr-style functions, starting with `~`, can be used in dplyr's `across()` function, such as `mutate(across(where(is.numeric), ~ .x ^ 2))`, and this function can only take one argument (`.x`)
4. True or False: If I have 10 tibbles I want to save to file, and they're all stored in a list, the best purrr function to use for saving these to file is `map()`.
5. True or False: Just as `c(c(1, 2), c(3, 4))` returns the vector of numbers from 1 to 4, `list(list(1, 2), list(3, 4))` returns the list of numbers from 1 to 4. 
6. True or False: `tibble(model = lm(mpg ~ wt, data = mtcars))` doesn't work because it doesn't use a `map` function.

