---
title: "Character Data"
output: html_document
---

From this topic, students are anticipated to be able to:

- Manipulate a character vector in R using the stringr package.
- Write simple regular expressions (regex).
- Apply stringr and regular expressions to manipulate data in tibbles.

## Resources

Video lecture:

- [Regular Expressions and stringr for Text Data](https://www.youtube.com/watch?v=I0dJ1zpxAtU) (only labelled as "age restricted" because it looks at real emails within the Enron company.)

Written material:

- Overview tutorials similar to our worksheet:
    - [stat545.com Chapter 11: character vectors](https://stat545.com/character-vectors.html)
    - [R4DS Chapter 15: strings](https://r4ds.hadley.nz/strings.html).
    - [R4DS Chapter 16: regular expressions](https://r4ds.hadley.nz/regexps).
- The stat545.com Chapter 11 on character vectors [has an elaborate discussion on useful resources](https://stat545.com/character-vectors.html#resources-3) for learning more about strings. 

## Strings 

You've used a bunch of strings at this point without knowing explicitly what they are: any time you surround text by `"`, you've been making a *string*: a storage format for text. In R, they are of type "character". 

```{r} 
sample_string <- "This is a string" 
typeof(sample_string)
```

Two places where you'll often want to manipulate these in data analysis: 

- Cleaning up column/variable names 
- Cleaning up character column values 

**Good to know**: Constructing strings out of characters and numbers is intuitive, but there's a gotcha involving particular symbols with special meaning in R. For example, try running `quote <- """` in R; it won't work, because the `"` symbol is interpreted as you trying to make a string! To literally include a quote in a string, you can use the `\` character to "escape" it: 

```{r} 
single_quote <- "\""
cat(single_quote)
```

You can see more examples of special characters and how to escape them in [R4DS Chapter 15.2](https://r4ds.hadley.nz/strings#creating-a-string). 

## Working with strings 

Our main tools for working with strings will be the powerful `stringr` package in the tidyverse paired with regular expressions. We think the best way to start learning these is through the guided tutorial in Worksheet B2.

## Test Your Understanding

Use these questions to check your understanding of the material.

1. True or False: The regular expression `[ab][ab]` will match "ab", "aa", and "bb" as possibilities, whereas `[ab]{2}` will only match "aa" or "bb".
2. True or False: The regular expression `[ab][ab]` will match "ab", "aa", and "bb" as possibilities, whereas `(ab)(ab)` will only match "aa" or "bb".
3. True or False: The regular expression `^ab` will match "ab" as the first characters to a string, whereas `[^ab]` will match "a" or "b" as being the first character to a string.

### Agenda 

## Class 1

- **Before class**, start working on parts I and II of Worksheet B2. 
- Class will be dedicated to getting your questions answered. 
- Done early? Then do the optional R4DS Strings and R4DS Regular expressions readings (linked above), and do exercises for extra practice. 

## Class 2
- **Before class**, start working on parts II and III of Worksheet B2. 
- Class will be dedicated to getting your questions answered. 
- Done early? Then do the optional R4DS Strings and R4DS Regular expressions readings (linked above), and do exercises for extra practice. Or, start Assignment B4. 

