---
title: "The Model-Fitting Paradigm in R"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(tidymodels)
library(gapminder)
```

From today's class, students are anticipated to be able to:

- make a model object in R, using `lm()` as an example.
- write a formula in R.
- predict on a model object with the `broom::augment()` and `predict()` functions.
- extract information from a model object using `broom::tidy()`, `broom::glance()`, and traditional means.

Note that there is a new tidyverse-like framework of packages to help with modelling. It's called [tidymodels](https://www.tidymodels.org/). 
<!-- Tidymodels has come a long way since it first started, and I can see it taking the place of this lecture in the future (or at least becoming a part of the lecture). -->

## Resources

Video lecture:

- [The Model-Fitting Paradigm in R](https://youtu.be/jQqCDeJYzao)

Other resources, in addition to the notes below:

- The [broom vignette](https://cran.r-project.org/web/packages/broom/vignettes/broom.html) is useful for learning the details of broom, a package for tidying the output of models.
- If you want a quick intro to linear regression with R, check out this [U Chicago Tutorial](https://cfss.uchicago.edu/notes/linear-models/) on model fitting in R (just the linear regression part).
- If you're eager to learn more about models in general, [An Introduction to Statistical Learning](https://www-bcf.usc.edu/~gareth/ISL/) is an approachable read (the book is freely available online)
- [Mike Marin's R playlist on YouTube](https://www.youtube.com/playlist?list=PLqzoL9-eJTNBlVXxWvJkq0dtVut2sICUW) helps you use R to obtain statistical results. 

# Model-fitting in R

Data wrangling and plotting can get you pretty far when it comes to drawing insight from your data. But, sometimes you need to go further. For example:

- Investigate the relationship between two or more variables
- Predict the outcome of a variable given information about other variables

These typically involve fitting models, as opposed to simple computations than can be done with tidyverse packages like dplyr.

This tutorial is _not_ about the specifics of fitting a model. Even though a few references to statistical concepts are made, just take these for face value. 

## Example: linear model

Here are a couple tasks that model-fitting would be useful to address. 

1. A car weighs 4000 lbs. Provide a numerical prediction on its mpg (miles per gallon). 
2. Does the weight of a car influence its mpg?
3. How many more miles per gallon can we expect of a car that weighs 1000 lbs less than another car?

A simple scatterplot will give us a general idea, but can't give us specifics. Here, we use the `mtcars` dataset in the `datasets` package. A linear model is one example of a model that can attempt to answer all three -- the line corresponding to the fitted model has been added to the scatterplot.

```{r, fig.width = 5, fig.height = 3, echo = FALSE, message = FALSE}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  labs(x = "Weight (1000's of lbs)") +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() 
fit <- lm(mpg ~ wt, data = mtcars)
```

(By the way, you can plot the model using `ggplot2`'s `geom_smooth()` layer -- but you can't interact with the model downstream.)

We can make a prediction of mpg for a car weighing 3,000 lbs by evaluating the line at X = 3:

```{r, echo = FALSE}
predict(fit, newdata = tibble(wt = 3)) %>% 
  unname()
```

We can find the p-value testing the null hypothesis that the true slope of the regression line is 0:

```{r, echo = FALSE}
glance(fit)$p.value %>% 
  unname()
```

And we can calculate how many more miles per gallon a car weighing 1000 less lbs would be expected to have through the slope:

```{r, echo = FALSE}
-tidy(fit)$estimate[2]
```


## Fitting a model in R

Fitting a model in R typically involves using a function in the following format:

```
method(formula, data, options)
```

__Method__:

A function such as:

- Linear Regression: `lm`
- Generalized Linear Regression: `glm`
- Local regression: `loess`
- Quantile regression: `quantreg::rq`
- ...

__Formula__:

In R, takes the form `y ~ x1 + x2 + ... + xp` (use column names in your data frame), where `y` here means the outcome variable you're interested in viewing in relation to other variables `x1`, `x2`, ...

__Data__: The data frame or tibble. Can omit, if variables in the formula are defined in environment

__Options__: Specific to the method, and include ways to customize the model.


Running the code returns an object -- usually a special type of list -- that you can then work with to extract results. 

For example, let's fit a linear regression model on a car's mileage per gallon (`mpg`, "Y" variable) from the car's weight (`wt`, "X" variable). Notice that no special options are needed for `lm()`.

```{r}
(my_lm <- lm(mpg ~ wt, data = mtcars))
```

Printing the model to the screen might lead you to incorrectly conclude that the model object `my_lm` only contains the above text. The reality is, `my_lm` contains a lot more, but special instructions have been given to R to only print out a special digested version of the object. This behaviour tends to be true with model objects in general, not just for `lm()`.

## Probing the model: Broom

Now that you have the model object, there are typically three ways in which it's useful to probe and use the model object. The `broom` package has three crown functions that make it easy to extract each piece of information by converting your model into a tibble:

1. `tidy`: extract statistical summaries about each "component" of the model.
2. `augment`: add columns to the original data frame containing predictions.
3. `glance`: extract statistical summaries about the model as a whole (a 1-row tibble).

### `tidy()`

Use the `tidy()` function for a statistical summary of each component of your model, where each component gets a row in the output tibble. For `lm()`, `tidy()` gives one row per regression coefficient (slope and intercept).

```{r}
tidy(my_lm)
```

`tidy()` only works if it makes sense to talk about model "components".

### `augment()`

Use the `augment()` function to make predictions on a dataset by augmenting predictions as a new column to your data. By default, `augment()` uses the dataset that was used to fit the model.

```{r}
augment(my_lm) %>% 
  print(n = 5)
```

We can also predict on new datasets. Here are predictions of `mpg` for cars weighing 3, 4, and 5 thousand lbs. 

```{r, warning = FALSE}
augment(my_lm, newdata = tibble(wt = 3:5))
```

Notice that only the "X" variables are needed in the input tibble (`wt`), and that since the "Y" variable (`mpg`) wasn't provided, `augment()` couldn't calculate anything besides a prediction.

### `glance()`

Use the `glance()` function to extract a summary of the model as a whole, in the form of a one-row tibble. This will give you information related to the model fit.

```{r}
glance(my_lm)
```


## Probing the Model: Traditional Method

In order for a model to work with the `broom` package, someone has to go out of their way to contribute to the `broom` package for that model. While this has happened for many common models, many others remain without `broom` compatibility.

Here's how to work with these model objects in that case.

### Prediction

If `broom::augment()` doesn't work, then the developer of the model almost surely made it so that the `predict()` function works (not part of the `broom` package). The `predict()` function typically takes the same format of the `augment()` function, but usually doesn't return a tibble.

Here are the first 5 predictions of mpg on the `my_lm` object, defaulting to predictions made on the original data:

```{r}
predict(my_lm) %>% 
  unname() %>% 
  head(5)
```

Here are the predictions of mpg made for cars with a weight of 3, 4, and 5 thousand lbs:

```{r}
predict(my_lm, newdata = tibble(wt = 3:5)) %>% 
  unname()
```

Checking the documentation of the `predict()` function _for your model_ isn't obvious, because the `predict()` function is a "generic" function. Your best bet is to append the model name after a dot. For example:

- For a model fit with `lm()`, try `?predict.lm`
- For a model fit with `rq()`, try `?predict.rq` (from the `quantreg` package)

If that doesn't work, just google it: `"Predict function for rq"`

### Functions designed for the model

The developer of the model probably made a suite of other functions to help you probe the model object. Check the documentation to find these.

For example, I can find the regression coefficients of a `lm()` object with `coef()`:

```{r}
coef(my_lm)
```

Standard error of the residuals with `sigma()`:

```{r}
sigma(my_lm)
```

Don't be surprised if there is no function to extract what you want. If that's the case, read on...

### Manually Extracting Information

If you can't extract model information from built-in functions like `coef()` or `sigma()`, you can manually dig in to the model object. In most cases, a model object is just a list in disguise! (Lists are discussed in STAT 545B). You can therefore extract information like you would with any other list.

To help figure out what's in the list, use the `names()` function. Here are all the entries in the `lm()` object we fit earlier:

```{r}
names(my_lm)
```

Want the degrees of freedom of the residuals? Just extract that entry:

```{r}
my_lm$df.residual
```

To complicate things further this might only be _part_ of the information made available in the model object! More info might be stored in _yet another_ list after applying the `summary()` function:

```{r}
summary(my_lm)
```

Like the original model object, `summary()` _looks_ like it returns a bunch of text -- but it's actually another list! Let's again use `names()` to get the components of this list:

```{r}
names(summary(my_lm))
```

Now I can get more pieces of information, like the adjusted R-squared value:

```{r}
summary(my_lm)$adj.r.squared
```


## Plotting models with `geom_smooth`

We can plot models (with one predictor/ X variable) using `ggplot2` through the `geom_smooth()` layer. Specifying `method="lm"` gives us the linear regression fit (but only visually -- very difficult to extract model components!):

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
    geom_point() +
    geom_smooth(method = "lm") 
```

Let's visualize some relationships in the `gapminder` dataset.

Let's inspect Zimbabwe, which has a unique behavior in the `lifeExp` and `year` relationship.

```{r}
gapminder_Zimbabwe <- gapminder %>% 
  filter(country == "Zimbabwe")

gapminder_Zimbabwe %>% 
  ggplot(aes(year, lifeExp)) + 
  geom_point()
```
Let's try fitting a linear model to this relationship
```{r}
ggplot(gapminder_Zimbabwe, aes(year,lifeExp)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)
```
Now we will try to fit a second degree polynomial and see what would that look like.

```{r}
ggplot(gapminder_Zimbabwe, aes(year,lifeExp)) +
  geom_point() +
  geom_smooth(method = "lm",
              formula = y ~ poly(x,2), 
              se = F)
```

## Summary

1. `function(formula, data, options)` - most models in R follow this structure.

2. `broom::augment()` - uses a fitted model to obtain predictions. Puts this in a new column in existing `tibble`. Equivalent base-r function is `predict()`.

3. `broom::tidy()` - used to extract statistical information on each component of a model. Equivalent is `coef(summary(lm_object))`.

4. `broom::glance()` - used to extract statistical summaries on the whole model. Always returns a 1-row `tibble`.

5. `geom_smooth()` - used to add geom_layer that shows a fitted line to the data. `method` and `formula` arguments can be used to customize model.

## Test Your Understanding

1. TRUE or FALSE: the output of `glance` on the above returns only 1 row because there is only 1 explanatory variable in the model.
2. TRUE or FALSE: the output of `broom::tidy()` package is more "tidy"  than `coef(summary())` because the output is a `tibble`
3. TRUE or FALSE: We need to use a modelling function, such as `lm`, before we can graph a fitted line with `geom_smooth`.
4. TRUE or FALSE: If I wanted to find the amount that `lifeExp` changes per `year`, then I need to use `broom::tidy()`

---

## Your turn: learning to fit models in R

We think the best way to learn the basics from here is to work through the last part of Worksheet A4. 

## In-class Schedule 

### First part

- Attempt questions on the last part of Worksheet A4
- Get answers about questions you got stuck on 

### Second part

While this is not a modelling class, let's get a sense of where modelling would fit into a real data analysis by working through [the final part of the FEV case study](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/notes/supp-a09.Rmd). 

### Attribution

Model fitting intro by Vincenzo Coia with input from Victor Yuan. FEV case study added later by Lucy. 