---
title: "Lecture Notes: Special objects in R: Dates, Times, and Factors"
subtitle: "Version 1.0.0"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```


## Learning Objectives

From today's class, students are anticipated to be able to:

- Recognize what the factor class is 
- Convert data in a dataframe to factors
- Use the tsibble package to convert to various date formats
- Plot with factors
- Reorder levels within factors

## Parallel Resources

This topic parallels the following resources. Take about 30 minutes to get acquainted with the following before coming to class:

- [STAT 545 Episode 10-A: Special Data Types in R: Dates, Times, and Factors](https://youtu.be/TTNfLtEUFDQ)

We suggest referring to this material when engaging with today's topic.

## Additional Resources

- The [stat545.com](https://stat545.com/factors-boss.html) chapter on Factors
- The [R4DS](https://r4ds.had.co.nz/dates-and-times.html) chapter on Dates and Times
- The `tsibble` [vignette](https://cran.rstudio.com/web/packages/tsibble/vignettes/intro-tsibble.html)
- The [forcats](https://forcats.tidyverse.org) package page and reference guide on page. 


## Quiz

1. The output of the following code is equivalent to a factor with the letters "a" to "f".

``` r
(abc <- factor(letters[1:3]))
#> [1] a b c
#> Levels: a b c
(def <- factor(letters[4:6]))
#> [1] d e f
#> Levels: d e f
c(abc, def)
```

2. The output of the following code is a date object.

```r

library(lubridate)

date <- ymd("2020-10-13")

dt <- ymd_hms("2020-10-13 09:30:00")

c(date, dt)

```

## Demonstration


```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(palmerpenguins)
library(gapminder)
```

Factors: demonstrate:

- "Canada" and "Other" with `fct_other()`
- Re-ordering of levels in plots:
  - Legend matches line plots
  - Categories with counts are ordered
  - Categories with distributions are ordered by something of your choosing.
- Bar plot of the total number of penguins surveyed from each species -- noting that 5 species are missing.


Tsibbles (did not cover)


