---
title: "R Functions for Data Analysis"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "asis"}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3, 
                      fig.align = "center", error = TRUE)
```

From this topic, students are expected to be able to:

- Start getting a sense of when to make a function in a data analysis (we will build on this next week).
- Workflow for building a function: start interactively, wrap it as a function. `return()`. Argument names. 
- Fortify a function: 
  - generalize the function and use smart defaults; `NA` handling, and ellipses package [https://ellipsis.r-lib.org/](https://ellipsis.r-lib.org/)
  - provide useful error messages; sidebar: `if` statements
  - Unit tests, and (sidebar) assertions
- Data masking in a function.
- Documenting a function

What we're _not_ covering (this week):

- Documenting functions with roxygen2 (R packages)
- S3 objects (and other object oriented styles) (R packages)


## Resources

Video lecture: 

- [R Functions for Data Analysis](https://youtu.be/3nDgR7l5Tps)

Written resources:

- Basic function syntax in R: https://swcarpentry.github.io/r-novice-inflammation/02-func-R/
- When to use functions in your data analysis:
    - [stat545.com Functions, Parts 1-3](https://stat545.com/functions-part1.html)
    - [R4DS functions chapter](https://r4ds.had.co.nz/functions.html)


## Functions: basics

A user-defined function needs three things: 

1. A name - self explanatory.
2. Arguments, to take in values from the user. 
3. Body, to define what code is executed when the function is run. 

```{r, eval=FALSE}
name <- function(arguments) { 
  body  
}
```

Once defined, you can call the function using the following syntax: 
```{r, eval=FALSE}
name(new_arguments)
```

Here is an example of a user-defined function in action: 
```{r}
roll_d10 <- function(num_dice) { 
    sum(sample(1:10, num_dice, replace=TRUE))
}

roll_d10(2)
```

Some topics:

- Curly brackets vs. no curly brackets.
- `return()` vs. no `return()`.
- Argument defaults vs. no argument defaults.
- Argument names as "dummy variables", independent of what's named in your workspace. 

## Functions: an example

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(gapminder)
library(testthat)
library(patchwork)
```

Sometimes it's useful to work with the _log_ of a variable with a positive skew, such as `gapminder`'s GDP per capita:

```{r}
gdpPercap_plot <- 
  ggplot(gapminder, aes(gdpPercap)) +
    geom_histogram() +
    theme_minimal()

log_gdpPercap_plot <- 
  ggplot(gapminder, aes(log(gdpPercap))) +
    geom_histogram() +
    theme_minimal()

gdpPercap_plot + log_gdpPercap_plot
```

And, it's sometimes useful to scale a variable to have mean 0 and variance 1. We'll do both:

```{r}
gapminder %>% 
  mutate(gdpPercap_scaled = (log(gdpPercap) - mean(log(gdpPercap))) / sd(log(gdpPercap)),
         pop_scaled = (log(pop) - mean(log(pop))) / sd(log(pop)))

gapminder %>% 
  mutate(gdpPercap = (log(gdpPercap) - mean(log(gdpPercap))) / 
           sd(log(gdpPercap)),
         pop = (log(pop) - mean(log(pop))) / sd(log(pop)))
```

Where might there be repetition in this analysis? Where __is__ there repetition?

- Doing this scaling over and over again throughout the analysis.
- Repetition across multiple variables, like `pop` and `lifeExp`
- `gdpPercap` appears three times.

Solve the problem of repetition by writing _functions_.

Turn the above into a function.

```{r}
my_scale <- function(x, remove_na = TRUE, ...) {
  logx <- log(x)
  (logx - mean(logx, na.rm = remove_na, ...)) / sd(logx, na.rm = remove_na)
}
my_scale(1:10)
```

Now we can use the function in our previous dplyr workflow:

```{r}
gapminder %>% 
  mutate(gdpPercap = my_scale(gdpPercap),
         pop = my_scale(pop))
```

Each function you create shouldn't do too much, but rather should do one thing well. Don't [make a gigantic `tablestat()` function](https://github.com/statdivlab/rigr/blob/1a6f6984a3f1fd4d3cc6a7dba8a213c041cde33f/R/tableStat.R) that attempts to do everything. Modularizing computation into smaller functions makes troubleshooting much easier. (Sorry, Scott!! It had to be said!!!!)

Let's take a moment to talk about the "..." argument before we move on. In short, it lets the user of your function pass in any number of additional arguments. Above, we used "..." to allow the user flexibility to take advantage of secret undocumented features of the `mean()` function. 

However, that exact same flexibility can cause you headaches down the road. The [tidyverse design guide](https://design.tidyverse.org/dots-position.html) goes over details and best practices of "..." usage that help keep you happy and the user safe. 

## Function arguments

Some common function argument additions:

- `na.rm`, an indication of whether you should remove NA values or not.
- `verbose`, useful if you have either a potentially long computation, or a computation where you may want to see more details (such as function inversion or optimization).

## Documenting your function

Documentation should include, at a minimum:

1. A description of what the function does, at a high level. 
2. A description of the arguments (input). What type of objects are they?
3. A description of the output. Describe the object that the function returns.

You should use [roxygen2 tags](https://roxygen2.r-lib.org/articles/rd-formatting.html) to document your function, placed immediately above the function definition. Although roxygen2 tags are designed for use when creating R packages, they provide a standardized way to document a function -- and make it easy for you to migrate your function to an R package if need be!

Here's an example of a custom function I use a lot, with roxygen2 documentation. 
```{r} 
#' Takes the l2-norm of a vector.
#'
#' @keywords internal
#'
#' @param x the vector to be normed
#'
#' @return Returns the l2-norm of x.
norm_vec <- function(x) {
  sqrt(sum(x^2))
}
```

## Testing your functions

`testthat` is a collection of functions developed by Hadley Wickham that makes unit testing easy for developers. You can read more about the structure of tests [here](https://r-pkgs.org/testing-basics.html#test-organisation). In a nutshell, tests are organized hierarchically: **expectations** are grouped into **tests**. Functions that start with `expect_` describe the expected result of a computation (e.g. Does it have the right class?) - these are expectations. Tests are created with `test_that()` and they group together multiple *expectations* to test the output of a function (at their simplest level). 

Let's test the `my_scale()` function we made earlier, first informally.

```{r}
my_scale(exp(c(1, 3)), verbose = TRUE)
my_scale(exp(c(1, 3, NA)))
my_scale(1:10)
my_scale(1:10, trim = 0.1, foobar = 3)
```

Now, write a formal test that doesn't require eyeballs on the code:

```{r}
test_that("Basic testing", {
  expect_equal(my_scale(exp(c(1, 3))), (c(1, 3) - 2) / sqrt(2))
  expect_equal(my_scale(exp(c(1, 3, NA))), (c(1, 3, NA) - 2) / sqrt(2))
  expect_error(my_scale(letters))
})
```

Testing your functions is a way of recognizing that your data analysis will evolve, and is a way of knowing which of your functions need updating if you make a breaking change in your analysis. 

## A return to dplyr scoped variants with `across()`

Let's try our function applied to multiple columns of the `gapminder` dataset using dplyr's `across()` function, reducing even more repeated code.

```{r}
gapminder %>% 
  mutate(across(c("gdpPercap", "pop"), my_scale))
```

## Function arguments without quotation marks: curly-curly

Sometimes your function needs to take variable names without quotation marks. For example, `select(gapminder, lifeExp)` does not put quotation marks around `lifeExp` (the reason being that `lifeExp` is like a variable in our workspace, if we were to include column names in our R Environment).

Making your own functions having this capability involves a workaround. For example, consider making a function to produce a quick scatterplot between two columns in a dataset:

```{r}
quick_scatter <- function(data, x, y) {
  ggplot(data, aes(x, y)) +
    geom_point()
}
```

The function won't work:

```{r, error = TRUE}
quick_scatter(gapminder, x = gdpPercap, y = lifeExp)
```

The reason is that R is looking for variables named `gdpPercap` and `lifeExp` in the workspace, and cannot find them. To fix the problem, we can change the function definition so that `x` and `y` are embraced within two curly brackets -- an operator called "curly curly": 

```{r}
quick_scatter <- function(data, x, y) {
  ggplot(data, aes({{ x }}, {{ y }})) +
    geom_point()
}
```

Now the function will work:

```{r, fig.width = 3, fig.height = 2}
quick_scatter(gapminder, x = gdpPercap, y = lifeExp)
```

But, you can only use curly-curly when passing your function's argument to another function _that's anticipating a variable name without brackets_. For example, consider subsetting a data frame by a column name using `[[`, where quotation marks are needed:

```{r}
mtcars[["wt"]]
```

Here's a function that subsets a data frame using `[[`, not using curly-curly:

```{r}
my_subset <- function(data, variable) {
    data[[variable]]
}
```

The function is anticipating quotation marks, and will fail without them:

```{r, error = TRUE}
my_subset(mtcars, "wt") 
my_subset(mtcars, wt)
```

Using curly-curly doesn't grant the `[[` subsetting any data masking powers -- data masking (i.e. the practice of allowing you to refer to variables in the "current" data frame without quotation marks or other frills) must already be in place, as they are in `dplyr` and `ggplot2` functions:

```{r}
my_subset <- function(data, variable) {
    data[[{{ variable }}]]
}
```

We get an error if we don't include quotation marks:

```{r, error = TRUE}
my_subset(mtcars, wt)
```

Useful hint: in the `dplyr` documentation, if you spy the words "data masking" or "tidy selection", then you will need to curly-curly your arguments when using those functions within your custom function.

## Test Your Understanding

Consider the following code:

```
x <- 1
f <- function(t) {
  x <- -1
  if (x > 0) return(t)
  x * t + 1
  x * t
}
f(1)
```

1. True or False: the output of `f(1)` is `-1`. 
2. True or False: after running the above code, `x` is `-1`. 
3. True or False: Curly-curly `{{}}` can be used whenever we want to refer to column names outright, such as when subsetting in base R with `[[`.
4. You've made a function `quick_scatter()` that you can pipe a tibble into, and outputs a scatterplot of two variables. So far, you're only using the function to explore the data for your own eyes (perhaps even putting your code directly in the console). True or False: at this point, you should stop and make sure your function is documented and has formal unit tests. 

## Attribution 

Most of these notes were originally compiled by previous iterations of the instructional staff, including Vincenzo Coia. 