---
title: 'FEV Case Study: Graphing, With Solutions'
author: "Lucy"
date: "2023-08-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Review 

We'll continue exploring the FEV data set from last week. Let's start by loading the data and required packages.

```{r, message = FALSE} 
library(rigr) 
library(dplyr)
library(ggplot2)

fev_tbl <- as_tibble(fev) %>% mutate(across(sex:smoke, ~ as.factor(.x)))

fev_tbl <- fev_tbl %>% mutate(height = height*2.54) %>% 
  rename(height_cm = height)
```

Last week, we found that the mean FEV in the smoking group was substantially higher than the average FEV in the non-smoking group. That is, the smokers appear to have *better* lung function than the non-smokers. 

We also had a theory as to why: the association between FEV and smoking status may be confounded by age/size, as bigger lungs can exhale more air (all else equal) than smaller lungs.

## Smoking and lung function: a slightly more nuanced look

Once you are done each of these exercises, compare your plots with a partner or other members of a small group. Discuss the rationale behind your plotting choices and key pieces of information you think a viewer should take away from the plots. 

### Exercise 1

Use plot(s) to compare the distribution of FEV in the smoking group and the distribution of FEV in the non-smoking group. 

```{r}
# FILL IN HERE 
```


### Exercise 2

Use plot(s) to compare the distribution of FEV in smokers and the distribution of FEV in non-smokers at different ages. 

```{r} 
# FILL IN HERE 
```


### Exercise 3

Use plot(s) to compare the distribution of FEV in smokers and the distribution of FEV in non-smokers at different ages and genders.

```{r} 
# FILL IN HERE 
```
