# (4) List Columns

--- LAST YEAR'S CONTENT BELOW ---

```{r}
suppressPackageStartupMessages(library(tidyverse))
```


## Today's Agenda

Today's lessons are:

1. Review of `purrr` and piping.
2. Parallel mapping
3. List columns
4. An analysis using both

## Resources

All are from [Jenny's `purrr` tutorial](https://jennybc.github.io/purrr-tutorial/). Specifically:
 
- Parallel mapping: [Jenny's "Specifying the function in map() + parallel mapping"](https://jennybc.github.io/purrr-tutorial/ls03_map-function-syntax.html#parallel_map)
- List columns in data frames; nesting: [Jenny's "List Columns"](https://jennybc.github.io/purrr-tutorial/ls13_list-columns.html).

The all-encompassing application near the bottom of the worksheet is from [Jenny's "Sample from groups, n varies by group"](https://jennybc.github.io/purrr-tutorial/ls12_different-sized-samples.html)


## Review

### `purrr`

`map(x, f, ...)` returns a list with elements:

- `f(x[[1]], ...)`
- `f(x[[2]], ...)`
- ...

`map_dbl`, `map_lgl`, etc to return a vector.

We can specify a pre-defined `f`, or write it on-the-fly, or another way that we didn't touch on last time. Example with "squaring" function: 

- `map(x, square)` where `square <- function(t) t^2`;
- `map(x, function(t) t^2)`; or
- `map(x, ~ (.x)^2)` (function variable is `.x` by convention).

### piping: `.`

We know that `a %>% b()` is the same as `b(a)`.

Want to refer to `a` _in addition_ to the first argument? Specify it as a `.`. Example:

Gotcha: 

Case 1: LHS (= left-hand side) __not__ put as first argument when `.` appears in RHS:

```{r}
log(8, base=2)
## is identical to...
2 %>% log(8, base=.)
```

Case 2: LHS __is still__ put as first argument, even when `.` appears in RHS:

```{r}
c(ncol(mtcars), nrow(mtcars))
## is NOT identical to...
1:10 %>% c(min(.), max(.))
```

Trick: Use `{}` to "absorb" the placement of LHS as first argument:

```{r}
1:10 %>% {c(min(.), max(.))}
```


### Worksheet

To get participation points for today, we'll be filling out the `cm104-exercise.Rmd` file.

- [Rmd](https://github.com/STAT545-UBC/Classroom/blob/master/notes/cm104-exercise.Rmd)
- [html](http://stat545.com/Classroom/notes/cm104-exercise.nb.html).

Add this to your participation repo. 
