---
title: "cm103 Worksheet"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
suppressPackageStartupMessages(library(tidyverse)) # Loads purrr, too
library(repurrrsive) # Contains data examples
library(listviewer) # For viewing JSON/lists interactively
```

## Resources

This week, we'll be drawing from [Jenny Bryan's `purrr` tutorial](https://jennybc.github.io/purrr-tutorial/). Specifically:

- The [`map` tutorial](https://jennybc.github.io/purrr-tutorial/ls01_map-name-position-shortcuts.html)
- The [GitHub users tutorial](https://jennybc.github.io/purrr-tutorial/ls02_map-extraction-advanced.html) is similar.

In addition:

- Do you feel that you need a refresher on lists and vectors in R? Check out the [Vectors and Lists](https://jennybc.github.io/purrr-tutorial/bk00_vectors-and-lists.html) part of the tutorial.
- Are you familiar with the `apply` family of functions in Base R? You might find Jenny's ["Relationship to base and plyr functions"](https://jennybc.github.io/purrr-tutorial/bk01_base-functions.html) great for bridging the gap to `purrr`.


## Review Vectors and Lists

Vectors: 

- hold multiple entries of the same type.
- coerced to the least-informative data type in the vector.
- subset with single square brackets

Lists:

- hold multiple entries of anything.
- no entries are coerced (as a result of being able to hold anything)
- subset with `[`, `[[`, or `$`. 

## Review Vectorization

= element-wise application of a function.

Examples:

```{r}
(1:10) ^ 2
```

```{r}
(1:10) * (1:2)
```

```{r}
commute <- c(40, 20, 35, 15)
person <- c("Parveen", "Leo", "Shawn", "Emmotions")
str_c(person, " takes ", commute, " minutes to commute to work.")
```

## `purrr`

`purrr` is great when vectorization does not apply! 

Particularly useful if your data is in JSON format. 

Example: 

1. Explore the `wesanderson` list (comes with `repurrrsive` package). Hint: `str()` might help. It's a list of vectors. 
2. Use what you know about R to write code that extracts a vector of the first elements of each vector contained in `wesanderson`.

```{r}
YOUR_CODE_HERE
```

`str()` is not always useful! Try checking the structure of `got_chars` (= Game of Thrones characters):

```{r}
str(got_chars)
```


## Exploring lists

1. `str()`: embrace `list.len` and `max.level`

```{r}
str(got_chars)
```

2. Interactive exploration: `View()` and `listviewer::jsonedit(..., mode = "view")`

```{r}
View(got_chars)
jsonedit(got_chars, mode="view")
```

3. Don't be afraid to check out a subset! `names()` comes in handy, too.

```{r}
str(got_chars[[1]])
names(got_chars[[1]])
```


## Exploring `purrr` fundamentals

Apply a function to each element in a list/vector with `map`.

General usage: `purrr::map(VECTOR_OR_LIST, YOUR_FUNCTION)`

Note:

- `map` always returns a list.
- `YOUR_FUNCTION` can return anything!

Toy example 1: without using vectorization, take the square root of the following vector:

```{r}
x <- 1:10
map(FILL_THIS_IN, FILL_THIS_IN)
```

Toy example 2 (functions on-the-fly): without using vectorization, square each component of `x`:

```{r}
map(FILL_THIS_IN, FILL_THIS_IN)
```

Want a vector to be returned? Must specify the `typeof()` of the vector. Use `map_dbl()` to specify an output vector of type "double" for the above (check out the documentation `?map` for the acceptable vector types):

```{r}
map_dbl(x, sqrt)
```

Does your function have other arguments? You can specify them afterwards in the ellipses (`...`).

```{r}
map_chr(x, str_c, "potato.", sep="-")
```

## Your Turn: `purrr` fundamentals

1. Let's retry the `wesanderson` example: use `purrr` to write code that extracts a vector of the first elements of each vector contained in `wesanderson`. Play around with the following, too:
    - Use `head` instead of writing your own function.
    - Try different `map` functions, even the "wrong" types.
    - Use the ` ``[`` ` function if you're feeling daring.

```{r}
YOUR_CODE_HERE
```

2. Check that each character's list entry in `got_chars` has the same names as everyone else (that is, list component names, not character names). Here's one way to do it:
    1. Use the `names` function.
    2. Then, bind the names together in a single character.
    3. Then, apply the `table()` function.

```{r}
YOUR_CODE_HERE
```


## Shortcut functions

We can do the subsetting much easier with these shortcuts: just replace function with either:

- index you'd like to subset by, or
- name you'd like to subset by.

```{r}
map_chr(wesanderson, 1) # %>% unname()
```


## Your turn: shortcut functions

1. For the `got_chars` data:

- What are the titles of each character? 
- Is a vector output appropriate here?
- Use a pipe.

Note: each character's list entry has a component named `titles` as the 9th entry.

```{r}
YOUR_CODE_HERE
```

2. For the `got_chars` data:

- Extract a list of the "name" and "born" data for each person.
    - Use the function ` ``[`` ` or `extract()` (from the `magrittr` package, does the same thing) function to do the subsetting
- What happens when we switch to `map_dfr` instead of `map`? How about `map_dfc`?

```{r}
desired_info <- c("name", "born")
```

Note: as Jenny says, it's always safer (from a programming perspective) to work with output that's more predictable. The following would be safer, and is still readable:

```{r}
got_chars %>% {
    tibble(
        name = map_chr(., "name"),
        born = map_chr(., "born")
    )
}
```

Note the curly braces "tricks" the object prior to the pipe from entering as the first argument to `tibble`, because ` ``{`` ` only returns the last line evaluated. In this case there are two: the above code is equivalent to

```{r}
got_chars %>% {
    .
    tibble(
        name = map_chr(., "name"),
        born = map_chr(., "born")
    )
}
```

