---
title: "Class Meeting 4 Worksheet"
output:
    html_notebook:
        toc: true
        theme: cerulean
        number_sections: true
editor_options: 
  chunk_output_type: inline
---

```{r}
suppressPackageStartupMessages(library(tidyverse))
library(repurrrsive)
```

# Resources

All are from [Jenny's `purrr` tutorial](https://jennybc.github.io/purrr-tutorial/). Specifically:
 
- Parallel mapping: [Jenny's "Specifying the function in map() + parallel mapping"](https://jennybc.github.io/purrr-tutorial/ls03_map-function-syntax.html#parallel_map)
- List columns in data frames; nesting: [Jenny's "List Columns"](https://jennybc.github.io/purrr-tutorial/ls13_list-columns.html).

The all-encompassing application near the bottom of this worksheet is from [Jenny's "Sample from groups, n varies by group"](https://jennybc.github.io/purrr-tutorial/ls12_different-sized-samples.html)



# Parallel Mapping

We're going to work with toy cases first before the more realistic data analytic tasks, because they are easier to learn.

Want to vectorize over two iterables? Use the `map2` family:

```{r}
a <- c(1,2,3)
b <- c(4,5,6)
map2(a, b, function(x, y) x*y)
```
```{r}
map2(a, b, ~ .x * .y)
```
```{r}
map2(a, b, `*`)
```
```{r}
map2_dbl(a, b, `*`)
```

More than 2? Use the `pmap` family:

```{r}
a <- c(1,2,3)
b <- c(4,5,6)
c <- c(7,8,9)
pmap(list(a, b, c), function(x, y, z) x*y*z)
```
```{r}
pmap(list(a, b, c), ~ ..1 * ..2 * ..3)
```
```{r}
pmap_dbl(list(a, b, c), ~ ..1 * ..2 * ..3)
```

## Your Turn

Using the following two vectors...

```{r}
commute <- c(10, 50, 35)
name <- c("Parveen", "Kayden", "Shawn")
```

use `map2_chr()` to come up with the following output in three ways:

```{r}
str_c(name, " takes ", commute, " minutes to get to work.")
```

1. By defining a function before feeding it into `map2()` -- call it `comm_fun`.

```{r}
comm_fun <- FILL_THIS_IN
map2(FILL_THIS_IN, FILL_THIS_IN, FILL_THIS_IN)
```

2. By defining a function "on the fly" within the `map2()` function.

```{r}
map2(name, commute, function(t, s) FILL_THIS_IN)
```

3. By defining a formula.

```{r}
map2(name, commute, ~ FILL_THIS_IN)
```


# List columns

## What are they?

A tibble can hold a list as a column, too:

```{r}
(listcol_tib <- tibble(
  a = c(1,2,3),
  b = list(1,2,3),
  c = list(sum, sqrt, str_c),
  d = list(x=1, y=sum, z=iris)
))
```

Printing to screen doesn't reveal the contents! `str()` helps here:

```{r}
str(listcol_tib)
```

Extract a list column in the same way as a vector column:

```{r}
print(listcol_tib$a)  # Vector
```
```{r}
print(listcol_tib$b)  # List
```

This is where `map()` comes in handy! Let's make a tibble using the `got_chars` data, with two columns: "name" and "aliases", where "aliases" is a list-column (remember that each character can have a number of aliases different than 1):

1. Pipe `got_chars` into `{` with `tibble()`.
2. Specify the columns with `purrr` mappings.

```{r}
got_alias <- got_chars %>% FILL_THIS_IN
```

Write the solution down carefully -- we'll be referring to `got_alias` later. 

## Making: Your Turn

Extract the aliases of Melisandre (a character from Game of Thrones) from the `got_alias` data frame we just made. Try two approaches:

Approach 1: Without piping

1. Make a list of aliases by extracting the list column in `got_alias`.
2. Set the names of this new list as the character names (from the other column of `got_chars`).
3. Subset the newly-named list to Melisandre.

```{r}
(alias_list <- FILL_THIS_IN)
names(alias_list) <- FILL_THIS_IN
SUBSET_HERE
```

Approach 2: With piping

1. Pipe `got_alias` into the `setNames()` function, to make a list of aliases, named after the person. Do you need `{` here?
2. Then, pipe that into a subsetting function to subset Melisandre.

```{r}
got_alias %>% FILL_THIS_IN
```

## Nesting/Unnesting; Operating

__Question__: What would tidy data of `got_alias` look like? Remember what `got_alias` holds:

```{r}
got_alias
```

Let's make a tidy data frame! First, let's take a closer look at `tidyr::unnest()` after making a tibble of preferred ice cream flavours:

```{r}
(icecream <- tibble(
    name = c("Jacob", "Elena", "Mitchell"),
    flav = list(c("strawberry", "chocolate", "lemon"),
                c("straciatella", "strawberry"),
                c("garlic", "tiger tail"))
))
```

I can make a tidy data frame _without_ list columns using `tidyr::unnest()`:

```{r}
icecream %>% 
    unnest(flav)
```

How would I subset all people that like strawberry ice cream? We can either use the tidy data, or the list data directly:

From "normal" tidy data:

```{r}
icecream %>% 
    unnest(flav) %>% 
    filter(flav == "strawberry")
```

From list-column data:

```{r}
icecream %>% 
  filter(map_lgl(flav, ~ FILL_THIS_IN))
```


## Nesting/Unnesting: Your Turn

`unnest()` the `got_alias` tibble. Hint: there should be a hiccup. Check out the `str()`ucture of `got_alias` -- are all elements of the list column vectors? Would using `tidyr::drop_na()` be a good idea here?

```{r}
FILL_THIS_IN
```

We can also do the opposite with `tidyr::nest()`. Try it with the `iris` data frame:

1. Group by species.
2. `nest()`!

```{r}
FILL_THIS_IN
```

Keep the nested `iris` data frame above going! Keep piping:

- Fit a linear regression model with `lm()` to `Sepal.Length ~ Sepal.Width`, separately for each species. 
    - Inspect, to see what's going on.
- Get the slope and intercept information by applying `broom::tidy()` to the output of `lm()`.
- `unnest` the outputted data frames from `broom::tidy()`.

# Application: Time remaining?

If time remains, here is a good exercise to put everything together. 

[Hilary Parker tweet](https://twitter.com/hspter/status/739886244692295680): "How do you sample from groups, with a different sample size for each group?"

[Solution by Jenny Bryan](https://jennybc.github.io/purrr-tutorial/ls12_different-sized-samples.html):

1. Nest by species.
2. Specify sample size for each group.
3. Do the subsampling of each group.

Let's give it a try:

```{r}

```


# Summary:

- tibbles can hold columns that are lists, too!
    - Useful for holding variable-length data.
    - Useful for holding unusual data (example: a probability density function)
    - Whereas `dplyr` maps vectors of length `n` to `n`, or `n` to `1`...
    - ...list columns allow us to map `n` to any general length `m` (example: regression on groups)
- `purrr` is a useful tool for operating on list-columns.
- `purrr` allows for parallel mapping of iterables (vectors/lists) with the `map2` and `pmap` families.
