---
title: "Team Troubleshooting Deliverable 3"
output: github_document
---

## Attributions

Thanks to Icíar Fernández-Boyano for writing most of this document, and Vincenzo Coia for his edits.
   
## Instructions

There are **10 errors** in this Rmd. Will you find them all? Errors are marked by ## ERROR HERE ## in the code chunk in which they are located. *There are no hidden errors in this worksheet*. Hint: You should be able to knit the Rmd if all code is working correctly...but this will only catch the errors in code, and not the errors in logic or instruction-following!

**Notes:**

+ **Each code chunk contains a maximum of one single error.**

+ Errors may be in the form of **broken code** (e.g. using `" "` inappropriately, which returns an error when running the code), but can also be **running code that does not follow the instructions** (e.g. the goal of the code was to filter the first 3 rows, but instead, the first 3 columns were selected). Read the Rmd commentary carefully to find these.

## Setup

Below, I load the packages that are required to run this worksheet.

```{r, include = FALSE}
library(tidyverse)
library(lubridate)
library(palmerpenguins)
```

### Exercise 1

By now, you should be familiar with tidy data. Here's a *messy* tibble with data regarding smoking habits in a small office. 

```{r 1.0}
(smoker <- tribble(
  ~smoker, ~male, ~female, ~`non-binary`,
  "yes", 21, 10, 1,
  "no", 20, 12, 2
))
```

To tidy it, I want to make the data **longer** and store it into a new variable called `smoker_tidy`. The variables/columns in the tidy version of the data are:

+ `gender` ("female", "male", "non-binary")
+ `smoker` ("yes", "no")
+ `count`, which is a non-negative integer representing the number of observations.

Remember that you should **always look at the output of your code** to make sure it's doing what you're expecting.

Also remember that names that contain non-standard characters, like dashes `-`, can be worked with by surrounding the name with backticks.

```{r 1.1-error1}
# ERROR HERE - 1
(smoker_tidy <- smoker %>%
  pivot_longer(c(yes, no), names_to = "gender", values_to = "count"))
```

### Exercise 2

Remember the `penguins` dataset? I want to explore the relationship between `flipper_length_mm` and `body_mass_g` to see if there is a linear relationship between the two. To do so, I will use `ggplot2` to plot my data as a scatterplot where `x = flipper_length_mm` and `y = body_mass_g`, and I will **add a regression line (straight line) showing the confidence interval** using `geom_smooth()`. 

Remember that there can only be **one error** in a single code chunk.

```{r 2.0-error2}
# ERROR HERE - 2
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_smooth() +
  geom_point(method = "lm", se = TRUE) +
  theme_minimal() +
  labs(x = "Flipper Length (mm)",
       y = "Body Mass (g)")
```

### Exercise 3

`lubridate` is a `tidyverse` package that makes dealing with dates & times in R easy. It has a series of functions for parsing dates of a known order. For example, `ymd()` will parse dates with a *year-month-day* order, and `dmy()` will parse dates with a *day-month-year* order. You can check `lubridate`'s documentation, including parsing & manipulating dates, and instants, by running `?lubridate`. You can try below:

```{r 3.0}
?lubridate
```

With that said - time to work with some dates! I have stored Rosalind Franklin's birth date under the variable `rs`, and I would like to parse it using **the appropriate `lubridate` date function**, given the order in which I have stored it.

```{r 3.1-error3}
# ERROR HERE - 3
rs <- "July 25, 1920"
dym(rs) 
```

### Exercise 4

Being able to parse dates is nice, but `lubridate` is also very useful when it comes to answering questions such as "How many seconds are in a day?" or "How long has it been since...?". For example, I want to calculate how many *hours* it has been since the first landing on the moon, until today.

```{r 4.0-error4}
# ERROR HERE - 4
date_landing <- mdy("July 20, 1969")
difftime(today(), date_landing, units = "days")
```

### Exercise 5

Now, how many *days* will it have been since the landing on the moon, **in a week from now**?

```{r 5.0-error5}
# ERROR HERE - 5
difftime(today(), date_landing + weeks(1), units = "days")
```

### Exercise 6

The `palmerpenguins` dataset contains a tidy version of the dataset, `penguins`, and an untidy version, `penguins_raw`. Now... remember factors? In the words of Jenny Bryan, "factors are the variable type that useRs love to hate." It is how we store truly categorical information in R; the values in a factor are called "levels". They are particularly useful when it comes to plotting data, and fitting models. Which is why you should know how to use them! `forcats` is the `tidyverse` package that deals with factors.

Let's have a quick look at the `penguins_raw` dataset. There are many ways to do this, but I use `str()` because it is a fast way to see the *class* of each of the columns within the dataset.

```{r 6.0}
str(penguins_raw)
```

Let's change the `Sex` and `Species` columns - currently encoded as character columns - into factors, and store the dataset with these changes into a variable `penguins_new`.

```{r 6.1}
penguins_new <- 
  penguins_raw %>%
  mutate(across(c(Sex, Species), as.factor))

levels(penguins_new$Species) # check output
```

Now, I noticed that the levels for the factor `Species` could be simplified. Therefore, I want to **recode the levels** as follows, once again saving these changes into the `penguins_new` dataset, more specifically into the same column `Species`:

+ "Adelie Penguin (Pygoscelis adeliae)" to "Adelie"
+ "Chinstrap penguin (Pygoscelis antarctica)" to "Chinstrap"
+ "Gentoo penguin (Pygoscelis papua)" to "Gentoo"

```{r 6.2-error6}
# ERROR HERE - 6
penguins_new <-
  penguins_new %>%
  mutate(Species = rename(Species,
                          "Adelie Penguin (Pygoscelis adeliae)" = "Adelie",
                          "Chinstrap penguin (Pygoscelis antarctica)" = "Chinstrap",
                          "Gentoo penguin (Pygoscelis papua)" = "Gentoo"))
levels(penguins_new$Species) # check output
```

### Exercise 7

So far, we've converted columns `Sex` and `Species` into factors. It is common practice to convert **all character columns** into factors. Let's try that below; again, saving our changes into the `penguins_new` dataset.

```{r 7.0}
penguins_new <-
  penguins_new %>%
  mutate(across(where(is_character), as_factor))
```

Say I want to create a subset of the `penguins_new` dataset, that *only* contains data for **female** penguins from the **Adelie** and **Gentoo** species. Save this subset into `penguins_subset` and ensure that you *drop all unused levels* from this new subset. Remember to *check the output* to ensure that the code is doing what you expect it to do!

```{r 7.1-error7}
# ERROR HERE - 7
penguins_subset <-
  penguins_new %>%
  filter(Species %in% c("Adelie", "Gentoo") & Sex == "FEMALE") %>%
  drop_na() 
```

### Exercise 8

In the remaining 3 exercises (8, 9, and 10), we will practice joins! First, let's create a couple of tibbles that we can practice `dplyr` joins with. This data is a subset of the `bechdel` movies dataset by [FiveThirtyEight](https://github.com/fivethirtyeight/data/tree/master/bechdel). 

```{r 8.0}
movies <- tibble::tribble(
                                         ~movie,     ~bechdel_binary,      ~genre,                       ~distributor,
                              "American Hustle",              "pass",     "drama",      "Sony Pictures Entertainment",
                                  "Pacific Rim",              "fail",    "action",             "Warner Bros Pictures",
                             "The Great Gatsby",              "fail",     "drama",            " Warner Bros Pictures",
                                     "Hannibal",              "good",  "thriller",         "MGM Distribution Company",
 "Harry Potter and the Deathly Hallows: Part 1",              "pass",   "fantasy",             "Warner Bros Pictures",
                      "The Blair Witch Project",              "pass",    "horror",            "Artisan Entertainment",
                      "The Amazing Spider-Man",              "fail",    "fantasy",          "Sony Pictures Releasing"
  )

distributors <- tibble::tribble(
                          ~distributor, ~yr_founded,
         "Sony Pictures Entertainment",        1998,
                "Warner Bros Pictures",        1923,
            "MGM Distribution Company",        1924
  )
```

Join `movie` and `distributor` so that all rows in `movie` where there are no matching values in `distributor` are kept.

```{r 8.1-error8}
# ERROR HERE - 8
right_join(movies, distributors) 
```

### Exercise 9

Join `movies` and `distributors` so that all rows from `distributors` and all columns from `movies` and `distributors` are kept.

```{r 9.0-error9}
# ERROR HERE - 9
left_join(movies, distributors)
```

### Exercise 10

Join `movies` and `distributors` so that all rows and all columns from both datasets are present.

```{r 10.0-error10}
# ERROR HERE - 10
inner_join(movies, distributors)
```


