---
title: "Demo 11-A: Tibble Joins"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
library(palmerpenguins)
library(tidyverse)
```

This is a quick demo of joining with the penguins tibble that we saw in class.

Recall the `penguins` tibble:

```{r}
penguins
```

There are only 3 islands:

```{r}
penguins %>% 
  pull(island) %>% 
  levels()
```

Suppose we think land mass of the islands has something to do with body mass. We'd want to add land mass to the tibble. Here's a tibble of (fake) land mass values for islands in the area (I added another fake one, too):

```{r}
(mass <- tribble(
  ~ island, ~ mass,
  "Biscoe"    , 500,
  "Dream"     , 600,
  "Foo Island", 300,
  "Torgersen" , 10000
))
```

Add them to the original tibble:

```{r}
(big <- left_join(penguins, mass, by = "island"))
```

By the way, `left_join()` is probably the most useful and the most used join (in my opinion).

Notice that Foo Island hasn't changed anything about our new tibble -- there are still three islands. Let's see: 

(Note: it seems `island` is no longer a factor, so we'll use `unique()` instead of `levels()` to get all unique islands):

```{r}
big %>% 
  pull(island) %>% 
  unique()
```

Now suppose you want to do the opposite: you only have one big tibble, and you want to store island land mass in a new tibble to save storage space on your computer. We have two options:

```{r}
big %>%
  group_by(island) %>%
  summarise(mass = unique(mass))
```


```{r}
big %>%
  group_by(island) %>%
  summarise(mass = mass[1])
```

The second example is less desirable: what if -- for whatever reason -- there's _more than one_ land mass for an island? `mass[1]` wouldn't catch that, but `unique()` would. Better to use `unique()` so that you'll be made aware that there's more than one land mass.

