---
title: "Demonstration 8-A: tidyr for tidy data"
output: html_document
---

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(palmerpenguins)
library(gapminder)
knitr::opts_chunk$set(error = TRUE)
guestlist <- read_csv(here::here("content", "data", "wedding", "attend.csv"))
```

## Review

### Collaborative Milestone 2

Logical operators: `%in%`, `!=`, `!`

```{r}
v <- c("foo", "baz", "foo", "bar")
```


- Avoid using deprecated functions: use `across()` instead of scoped variants like `mutate_at()` and `summarise_if()`.
- Troubleshooting? Check the function's documentation.
  - Easy to access with `?function`
  - Typically concise. Not like those old fashioned "User Manuals" [like this WinBUGS manual](https://www.mrc-bsu.cam.ac.uk/wp-content/uploads/manual14.pdf).

### Worksheet 3-A

Any requests? If not, Question 1.3:

> In a single expression, make a tibble with the following columns for each island in the penguins data set:
> 
> - What is the mean of each numeric variable in the penguins dataset in each island? Keep the column names the same.
> - How many penguins are there in each island? Add this to a column named `n`.

```{r}

```


## tidyr Demo

Let's work with this guestlist dataset.

```{r}
guestlist
```

### 1. Country relationship

Using the `gapminder` dataset: What's the relationship between Canada's GDP per capita and the United Kingdom's? First, identify the tidy format.

```{r}
gapminder %>% 
  FILL_THIS_IN
```

### 2. Who's PENDING?

Our task: extract all "PENDING" entries, paired with who it is, and what information is pending.

What is a tidy data format for this question? Do the tidying.

```{r}
guestlist %>% 
  FILL_THIS_IN
```

### Quiz

1. True or False: it's possible to lengthen a tibble "all the way" so that it can't be lengthened anymore.
2. True or False: To make a scatterplot viewing the relationship of Africa's GDP per capita vs. Europe's, use `pivot_wider()`. 


### 3. Relationship between PENDING

What's the relationship between whether or not a meal is PENDING, and the attendance status? First determine what a tidy format is for this question, then answer the question. 

```{r}
guestlist %>% 
  FILL_THIS_IN
```

### 4. Split name

Split the guest name into first and last name.

```{r}
guestlist %>% 
  FILL_THIS_IN
```




## Deprecated functions: spread and gather

Pivoting used to be achieved through the tidyr `spread()` and `gather()` functions. These are deprecated and should not be used. 
