---
title: "Lecture 13-A: Choose Your Adventure"
output: html_document
---

```{r}
library(tidyverse)
library(gapminder)
library(palmerpenguins)
knitr::opts_chunk$set(error = TRUE)
```


# Part I: R Basics beyond the worksheet

## Curly brackets

Curly brackets can hold multiple lines of code, and outputs the result at the end.

There's not a whole lot of difference if you just using them on their own:

```{r}
{
  ADD_SOME_CODE_HERE
}
```

They start to matter when you need to add multiple lines of code as an argument.

### Question 1

You've encountered an error in the following code, and want to ask for help. Write a reprex to send over to someone, and paste it below the code chunk.

```
library(tidyverse)
library(gapminder)
gapminder %>% 
   group_by(Country) %>% 
   summarise(mean_pop == mean(pop))
```

```{r}
reprex::reprex(FILL_THIS_IN)
```

### Question 2

Sample 10 rows randomly from the `gapminder` tibble using `dplyr::sample_n()`, so that you get the same results each time. Do this all in a single pipe. 

```{r}
gapminder %>% 
  sample_n(10)
```


## Attributes

Did you know that a tibble is actually just a list?

```{r}
is.list(penguins)
```

Why, then, doesn't it look like a list? Because it carries metadata that tells R to treat it in a special way. The metadata is held in something called _attributes_.

The attributes of `penguins`:

```{r}

```

Remove attributes with `unclass()`:

```{r}

```

### Question 3

Make the following list stored in `homemade_tibble` a tibble by changing the attributes.

```{r}
(homemade_tibble <- list(
  species = c("Chinstrap", "Adelie"),
  body_mass_g = c(3405, 2304)
))
```

```{r}
attr(FILL_THIS_IN, which = "class") <- FILL_THIS_IN
attr(FILL_THIS_IN, which = "row.names") <- FILL_THIS_IN
homemade_tibble
```

### Question 4

Here's a regression model:

```{r}
(my_model <- lm(mpg ~ wt, data = mtcars))
```

What object is this? Use `unclass()` to find out. What are its attributes?

```{r}

```

## Style Guide

Write code that's easier to read by referring to the [tidyverse style guide](https://style.tidyverse.org/)

# Part 2: Turning Your GitHub Repo into a Website

It's common to turn your GitHub repository into a website. Case in point: stat545.stat.ubc.ca; vincenzocoia.com. 
At the very least, this is useful for making your html files visible as a webpage.


## GitHub Pages

Activate GitHub pages in your repo's settings, and your repository's files will be served on a website. 

By default, your main README is shown on the website homepage. So, if you put links to other files in your README, you now have a navigable (albeit rudimentary) website.

If you want a beautiful website, you have a couple options:

1. Code up your own css files etc. (Hard)
2. Use a template (easier)
  1. GitHub Pages is powered by [Jekyll](https://jekyllrb.com/docs/github-pages/) behind the scenes. Dean Attali made a [template called Beautiful Jekyll](https://github.com/daattali/beautiful-jekyll). I used this for some time.
  2. Use blogdown (see the section after next). stat545.stat.ubc.ca and vincenzocoia.com both use this.

## Viewing an HTML file from your GitHub repo

Viewing an HTML file on GitHub only shows you the HTML code, not the rendered script. You'll need to provide a link to a rendered, viewable version of each HTML file you produce. Here's how you do that using [_GitHub Pages_](https://pages.github.com/):

1. Enable "GitHub pages" on your repo:
    - Go to "settings" on your repo, and stay on the default "Options" tab.
    - Scroll down to the "GitHub Pages" section.
    - Under "Source", click the "None" drop-down button, and select the branch you want to turn into a website (probably "Master").
2. Also under the "GitHub Pages" section, you'll find your website URL. Make note of this URL.
	- This URL will show a rendered version of your repo's README, but this is not important.

Just by enabling GitHub Pages, your HTML files are now viewable rendered by your browser. Technically, you just turned your GitHub repo into a website. The only trick is _navigating_ to that HTML page, since there's no default _interface_ to your website.

3. Obtain the URL to  HTML file on your repo:
    - Start with your GitHub Pages URL that you made note of in Step 2. This points to the root of your repository. 
    - Get the path to an HTML file you want to view. It should be something like `/path/to/file.html` (in this case, `file.html` lives in the `to` folder, which lives in the `path` folder in the root of your repo). 
    - Append the path to the HTML file to your GitHub Pages URL. Try the URL to see that it works.
4. Make it easy for a visitor to your repo to find the rendered HTML file! Add a link to the rendered HTML file somewhere in your repo, probably in a README file in your homework folder.

## Blogdown

Blogdown (an R package) is actually just a wrapper around a non-R software called hugo. 

Let's explore:

1. Check out the [blogdown website](https://bookdown.org/yihui/blogdown/). This is really the only step you need.
2. Check out the [recommended workflow](https://bookdown.org/yihui/blogdown/workflow.html)
3. Add this updated html page to the STAT 545 website?

# Part 3: Making Presentation Slides from Rmd

1. Try knitting this to `ioslides_presentation`.
2. Take a look at [other options](https://rmarkdown.rstudio.com/lesson-11.html) beside ioslides.
  - Feeling fancy? Try `output: revealjs::revealjs_presentation`.