---
title: "Demonstration 7-A: dplyr Part 2"
output: html_document
---

_Sept 29, 2020_

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(gapminder)
gapminder
```


### Demo 1

Compute a weighted mean of life expectancy for different years, weighting by population.

```{r}
gapminder %>% 
  group_by(year) %>% 
  mutate(wmean = sum(lifeExp * pop) / sum(pop))
```

What if I wanted those values repeated for each line?


### Demo 2

For each continent and year, how many countries have a "low" life expectancy? Let's say life expectancy is "low" if it's less than the median most recent life expectancy for each continent. 

```{r}
v <- c(4, 7, 2, 4, 5, 9)
v[v < 6]
v < 6
v[1:2]

gapminder %>% 
  group_by(continent) %>% 
  mutate(median = median(lifeExp[year == 2007])) %>% 
  filter(lifeExp < median) %>% 
  count(continent, year)
  # group_by(continent, year) %>% 
  # summarise(n = n())

gapminder %>% 
  mutate(asia = continent == "Asia")
```

### Quiz

True or False?

1. `summarise()` reduces each group down to one row, and `mutate()` preserves the number of rows. But, between the two, there's a way to reduce each group down to _two_ rows.
2. `across()` is useful, but it doesn't allow you to make new column names. 


### Demo 3

Get the maximum and minimum `gdpPercap` for all continents.

```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarise(max_g = max(gdpPercap),
            min_g = min(gdpPercap)) %>% 
  group_by(continent == "Asia") %>% 
  summarise(max2 = max(max_g))
```


### Demo 4

Filter gapminder to all entries that have experienced a drop in life expectancy. Hint: `dplyr::lag()`.

```{r}
v <- 1:4
v
dplyr::lag(v)

gapminder %>% 
  group_by(country) %>% 
  arrange(year) %>% 
  mutate(inc = lifeExp - lag(lifeExp)) %>% 
  filter(inc < 0)
```


### Demo 5

Filter gapminder to contain six rows: the rows with the three largest GDP per capita, and the rows with the three smallest GDP per capita. Don't make any excerpts.

```{r}
gap1 <- filter(gapminder, continent == "Europe")
gap2 <- summarise(gap1, mean_gdp = mean(gdpPercap))
gap2

nrow(gapminder)

gapminder %>% 
  arrange(gdpPercap) %>% 
  slice(c(1   , 2   , 3, 1702, 1703 ,1704)
    #c(1:3, (n() - 2):n())
  )

c(1:3, (1704 - 2):1704)
c(1:3, 1702:1704)


gapminder %>% 
  group_by(continent) %>% 
  mutate(n_function = n())
```


```{r}
gapminder %>% 
  mutate(decade = floor(year / 10)) %>% 
  group_by(country, continent, decade) %>% 
  summarise(across(where(is.numeric), mean)) %>% 
  select(-year)
```


