---
title: 'Demonstration 9-A: Model Fitting Paradigm'
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Demo

Topics for today's demo:

- fitting a linear regression model using `lm`
- how to extract modelling information using the package `broom`
- plotting models using `geom_smooth`

```{r}
library(tidyverse)
library(gapminder)
library(broom)
```

# What is Model-Fitting? 

A way to test if there is a "relationship" between two or more variables

Let's look at dataset `mtcars` as an example:

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(x = "Weight (1000's of lbs)")
```

1. Is there a relationship between `wt` (car weight) and `mpg` (fuel efficiency)?

2. Or another way to ask, does car weight have some information that tells us about fuel efficiency?

Modelling is a way to quantify this relationship. 

# Fitting a model in R

Most model fitting functions have a common format:

```
function(formula, data, options)
```

1. **Formula** 

- describes the relationship you want to test (e.g. `wt ~ mpg`)
- Formulas look like this: `y ~ x1 + x2 + ...`

2. **data** 

- the `data.frame`/`tibble` that contains your variables
- can omit, if variables in formula are defined in environment

e.g. 
```{r}
y1 <- c(1,2,3)
x1 <- c(0.1, 0.2, 0.3)
lm(y1~x1)

# same as
d1 <- tibble(y1 = y1, x1 = x1) # store y1 and x2 into tibble
rm(y1, x1)                     # remove y1 and x2 from environment
lm(y1~x1, data = d1)           
```


3. **options** 

- ways to customize the model
- specific to each type of model

The output of these is usually a **special _list_** object.

```{r}
my_model <- lm(y1~x1, data = d1) 
str(my_model)
```

Some model functions in R:

- Linear Regression: `lm`
- Generalized Linear Regression: `glm`
- Local regression: `loess`
- Quantile regression: `quantreg::rq`
- ...

Let's fit a linear model with between `wt` and `mpg` in the `mtcars` dataset:

```{r}
(my_lm <- lm(mpg ~ wt, data = mtcars))
```


# `predict()`

**Functions:** `broom::augment()` and `predict()`

We can use the `predict()` or `broom::augment()` to make predictions from the model (default is to use fitting/training data). 

```{r predict-training}
predict(my_lm) %>% 
  head()
```


```{r augment-training}
augment(my_lm)
```


Or we can predict on a new dataset:

```{r predict-newdata}
wt1 <- tibble(wt = seq(18:23))
wt1

predict(my_lm,wt1)
```

```{r augment-newdata}
augment(my_lm, newdata = wt1)

wt1 %>%
  mutate(mpg = c(32, 26, 20, 14, 8, 2)) %>%
  augment(my_lm, newdata = .)
```
# `summary()`

`lm` output lacks information:

```{r}
my_lm
```

To get more information, we can use `summary`:

```{r}
summary(my_lm)
```

`summary` is useful for reading, but it's not easy to work with this output

# Introduce `broom::tidy`

- extracts statistical summaries for each component of the model
- returns in tabular format
- useful for _interpretation_

```{r}
tidy(my_lm)
```

```{r}
coef(summary(my_lm)) 
```

# `broom::glance`

- gives statistical summary on model as a whole
- useful for checking goodness of fit

```{r}
glance(my_lm)
```

# Quiz 1

- TRUE or FALSE: the output of `glance` on the above returns only 1 row because there is only 1 explanatory variable in the model.
- TRUE or FALSE: the output of `broom::tidy()` package is more "tidy"  than `coef(summary())` because the output is a `tibble`

# Plotting linear models with `geom_smooth`

We can plot models (with one predictor/ X variable) using `ggplot2` through the `geom_smooth()` layer. Specifying `method="lm"` gives us the linear regression fit (but only visually!):

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
    #geom_point() +
    geom_smooth(method="lm") 
```

Let's visualize some relationships in the `gapminder` dataset.

Lets the country "Zimbabwe", which has a unique behavior in the `lifeExp` and `year` relationship.

```{r}
gapminder_Zimbabwe <- gapminder %>% 
  filter(country == "Zimbabwe")

gapminder_Zimbabwe %>% 
  ggplot(aes(year, lifeExp)) + 
  geom_point()
```
Let's try fitting a linear model to this relationship
```{r}
ggplot(gapminder_Zimbabwe, aes(year,lifeExp)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)
```
Now we will try to fit a second degree polynomial and see what would that look like.

```{r}
ggplot(gapminder_Zimbabwe, aes(year,lifeExp)) +
  geom_point() +
  geom_smooth(method = "lm",
              formula = y~poly(x,2), 
              se = F)
#?geom_smooth
```

# Quiz 2

1. TRUE or FALSE: We need to use a modelling function, such as `lm`, before we can graph a fitted line with `geom_smooth`.
2. TRUE or FALSE: If I wanted to find the amount that `lifeExp` changes per `year`, then I need to use `broom::tidy()`

# Summary

1. `function(formula, data, options)` - most models in R follow this structure.

2. `broom::augment()` - uses a fitted model to obtain predictions. Puts this in a new column in existing `tibble`. Equivalent base-r function is `predict()`.

3. `broom::tidy()` - used to extract statistical information on each component of a model. Equivalent is `coef(summary(lm_object))`.

4. `broom::glance()` - used to extract statistical summaries on the whole model. Always returns a 1-row `tibble`.

5. `geom_smooth()` - used to add geom_layer that shows a fitted line to the data. `method` and `formula` arguments can be used to customize model.