---
title: "assignment_01"
author: "Firas Moosvi"
date: "Updated `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 01

Below are 6 exercises to help you practice vectorizing for-loops, and converting for-loops into vectorized functions, here using `purr:map()`.
This assignment was designed to be completed in about 60 minutes and is deliberately designed to be shorter to accomodate the team project.

```{r message = FALSE, warning = FALSE}
library(tidyverse)
mtcars_original <- mtcars
```
### Task 1: Convert for-loop code to vectorized functions using purrr::map()

#### 1.1 - Below is sample code that computes the mean of every column in the `mtcars` dataset. Use the appropriate `purrr::map()` function to vectorize this task.

```{r prob1}
output <- vector("double", ncol(mtcars))
for (c in seq_along(mtcars)){
  output[[c]] <- mean(mtcars[[c]])
}
output
```

Your solution to vectorize the code above (using `purrr::map()`):

```{r soln1}
# add solution here
output
```

#### 1.2 - Below is sample code that divides the values in each column of the `mtcars` dataset by the maximum in that column. Use the appropriate `purrr::map()` function to vectorize this task.

```{r prob2}
for (c in seq_along(mtcars)){
  mtcars[[c]] <- mtcars[[c]] / max(mtcars[[c]], na.rm = TRUE)
}
head(mtcars)
```

Your solution to vectorize the code above (using `purrr::map()`):

```{r soln2}
# add solution here
```
#### 1.3 - Below is sample code creates a ggplot with fuel efficiency on the x-axis and horsepower on the y-axis for three cylinder levels (4, 6, 8). Use the appropriate `purrr::map()` function to vectorize this task.

```{r prob3}
cylinders <- sort(unique(mtcars_original[['cyl']]))
plots <- vector("list",length(cylinders))

for (d in 1:length(cylinders)){
  plots[[d]] <- mtcars_original %>% 
    filter(cyl == cylinders[d]) %>%
    ggplot() + 
    theme_bw() +
    geom_point(aes(x = mpg, y = hp) ) + 
    labs(x = 'Fuel efficiency (mpg)',
         y = 'Horsepower (hp)') + 
    ggtitle(glue::glue("Horsepower and Fuel efficiency for ", cylinders[d], " cylinders"))
}
plots
```

Your solution to vectorize the code above (using `purrr::map()`):

```{r soln3}
# add solution here
```

### Task 2: Convert the vectorized tasks into one that uses for-loops

#### 2.1  - Below is sample code that computes the number of unique values in each column of `mtcars` in a vectorized way. Turn this vectorized function into one that uses a for loop that reproduces the output (numbers should be the same, the format can be different).
```{r prob4}
map(mtcars_original, unique) %>% 
  map_dbl(length)
```

Your solution to "un-vectorize" the code above for loops:

```{r soln4}
# add solution here
```

#### 2.2 - Below is sample code that takes input from various columns in the `diamonds` dataset and outputs a string containing information from the input. Use the appropriate `purrr::map()` function to vectorize this task.

*Hint:* Note the use of `map2` to iterate over two columns (price and carat).

```{r prob5}
dmonds <- diamonds %>% 
  slice(1:4)
output <- map2_chr(dmonds$price, dmonds$carat, 
                   ~paste0("Diamond #", nrow(dmonds) ,
                           " sold for $",.x,
                           " and was ",.y, " carats"))
```

Your solution to "un-vectorize" the code above for loops:

```{r soln5}
# add solution here
output
```

#### 2.3 - Below is sample code that splits the iris dataset into three species, and saves each dataframe as a separate file. Turn this vectorized function into one that uses a for loop that reproduces the output.

*Hint 1*: `map2(x, y, f)` is the same as `pmap(list(x, y), f)`

*Hint 2*: Multiple files can be saved in a vectorized way using the`pmap` function. The [`nest` function](https://tidyr.tidyverse.org/reference/nest.html) is used to create a list-column of data frames for each species from the `iris` dataset.

```{r prob6}
iris %>%
  nest(data = -Species) %>%
  pmap(~write_csv(.y, path = glue::glue(.x, ".csv")))
```

Your solution to "un-vectorize" the code above for loops:

```{r soln6}
# add solution here
```

## Additional Resources

1. Chapter 21 of [R for Data Science](https://r4ds.had.co.nz/iteration.html).
2. [Learn to purr](http://www.rebeccabarter.com/blog/2019-08-19_purrr/) blog post.
3. Chapter 9 of [Advanced R for Data Science](https://adv-r.hadley.nz/functionals.html).