---
title: "assignment_01"
author: "Firas Moosvi"
date: "Updated `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 01

```{r}
library(tidyverse)
```
### Task 1: Convert for-loop code to vectorized functions using purr:map()

1.1 - Below is sample code that computes the mean of every column in the `mtcars` dataset. Use the appropriate purr::map() function to vectorize this task.

```{r prob1}
output <- vector("double", ncol(mtcars))
for (c in seq_along(mtcars)){
  output[[c]] <- mean(mtcars[[c]])
}
output
```

Your solution to vectorize the code above (using `purr::map()`):

```{r soln1}
# add solution for loop 1 here
# BEGIN SOLUTION
output <- map(mtcars,mean)
# END SOLUTION
output
```

1.2 - Below is sample code that divides the values in each column of the `mtcars` dataset by the maximum in that column.

```{r prob2}
# add for loop 2 here
for (c in seq_along(mtcars)){
  mtcars[[c]] <- mtcars[[c]] / max(mtcars[[c]], na.rm = TRUE)
}
mtcars
```

```{r soln2}
# add solution for loop 2 here

```

iterate over two lists/dfs
```{r prob3}
# add for loop 3 here

```

```{r soln3}
# add solution for loop 3 here

```

### Task 2: Convert 3 functions into for-loops

```{r prob4}
# add for func 1 here

```

```{r soln4}
# add solution for func 1 here

```


```{r prob5}
# add for func 2 here

```

```{r soln5}
# add solution for func 2 here

```


```{r prob6}
# add for func 3 here

```

```{r soln6}
# add solution for func 3 here

```